# Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lingglosses)
```

Here I'm going to include an overview of what data this poster is based on, including information about my coding.

## Afar
### Details
Glottocode: afar1241 <br> 
Family: Cushitic <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@kamil2015]

(@afar1) @kamil2015, 291, example 907
```{r, echo=FALSE}
gloss_example(transliteration = "taamiit -íyye",
             glosses = "work.DIM say.3M.PFV",
             free_translation = "He worked a little.")
```

(@afar2) @kamil2015, 292, example 909
```{r, echo=FALSE}
gloss_example(transliteration = "boddiná mukuɖ -ítta",
             glosses = "tooth wobble say.3F.IPFV",
             free_translation = "The tooth moves.")
```

## Aimele
### Details
Glottocode: aime1238 <br> 
Family: Aimele <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Neither of the constituent verbs in the SVC in (@aimele1) is morphologically-marked as subordinate to a final verb. Instead, each verb is uninflected, except for the final verb, which bears predicate markers such as tense, aspect, mood, modality, and evidentiality. 
[@aiton2016, 201]

### Examples

(@aimele1) @aiton2016, 200, example 27
```{r, echo=FALSE}
gloss_example("koodulu suloobo -moo wɛlɛ sɛɛ -jaa",
             "downward {NAME} DAT shout say PST",
             "I shouted it down to Soloobu.")
```

(@aimele2) @aiton2016, 201, example 30
```{r, echo=FALSE}
gloss_example(transliteration = "nɛ: ti:ɸɛ: o:ga: ɛja: o:gɛ: di a:nɛ:",
              glosses = "1:SG afterwards pandanus seedling:ABS carry_in_bilum take go:PST",
              free_translation = "I went after him, taking the pandanus seeds in a bilum.")
```

(@aimele3) @aiton2016, 201, example 31
```{r, echo=FALSE}
gloss_example(transliteration = "sugu:lu: ɛna: hɛna: do:ga: dijɛ:=bi: kɛi",
               glosses = "school that:ABS go house:ABS build=IMP ASSER",
               free_translation = "Go build that building for the school!")
```

## Ami
### Details
Glottocode: ammi1238 <br> 
Family: Western Daly <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: ?

> Phonological liaison occurs between the two verbs... There are grounds here for proposing that each serial construction constitutes a single phonological conglomerate. There is no pause between the VPs and the single intonation contour points to their constituting a single clause. [@ford1998, 316]

### Examples

(@ami1) @ford1998, 316, example 8-54
```{r, echo=FALSE}
gloss_example(transliteration = "gama vulhut gama",
             glosses = "3MIN.S.R.stand beeswax 3MIN.S.R.stand",
             free_translation = "The beeswax is standing up.",
             comment = "['kamaɸ̀ʊl̪ʊkàma]")
```

(@ami2) @ford1998, 316, example 8-55
```{r, echo=FALSE}
gloss_example(transliteration = "guman-pitji gaya",
             glosses = "3MIN.A.R.poke-roll 3MIN.S.R.lie",
             free_translation = "He is still making fire.",
             comment = "['kʊmanp̀ɪcɪɣàja]")
```


## Ashéninka Perené
### Details
Glottocode: ashe1272 <br> 
Family: Arawakan <br>
Macroarea: South American

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

> Serialized predicates do not have more than two verbs in a verb sequence which are fully inflected grammatical and phonological words, capable of being integrated into discourse. There is no pause separating the verbs from each other... the serialized verbs are independent phonological words, each inflected for person, aspect, realis status, and mode. [@mihas2015, 163]

### Examples

> Citation [@mihas2015, 159, 163-172]

(@asheninka1) @mihas2015, 165, example 6.13.b
```{r, echo=FALSE}
gloss_example(transliteration = "o- ja -t -atz -i o- shimaa -t -a",
             glosses = "3NM.S go EP PROG REAL 3NM.S fish EP REAL",
             free_translation = "She went to fish.")
```

(@asheninka2) @mihas2015, 168, example 6.19.b
```{r, echo=FALSE}
gloss_example(transliteration = "p- aako -t -ashi -t -a pi- ma -atz -i -tsi",
             glosses = "2S overtake EP APL.INT EP REAL 2S sleep PROG REAL NEG",
             free_translation = "Don't stay asleep.")
```



## Baure
### Details
Glottocode: baur1253 <br> 
Family: Arawakan <br>
Macroarea: South America

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: ?

> Finite verbs with possible person cross-reference on the verb, as in Baure, may also occur in serial verb constructions. ... The coordination of predicates is identical with the coordination of clauses without connector. Serial verb constructions, on the other hand, are a number of verbs that are uttered in one clause. In Baure the two categories seem pretty well distinguishable... [@danielsen2007, 427]

### Examples

(@baure1) @danielsen2007, 428, example 144
```{r, echo=FALSE}
gloss_example(transliteration = "ro=kač-po-wo ro=pino-po",
             glosses = "3SG.M=go-PRFLX-COP 3SG.M=flee-PRFLX",
             annotation = "*rokačpow* *ropinop*",
             free_translation = "he (went away and) fled")
```





## Bardi
### Details
Glottocode: bard1255 <br> 
Family: Nyulnyulan <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1?

> [Serial verbs] occur in a single intonation contour (although there are also examples with breaks, and examples where intonation units and syntactic units are not isomorphic). 
[@bowern2013, 663-664]

> A further piece of evidence shows that these clauses are joined in a single higher structure rather than chained without embedding. It is possible to extract constituents from a lower clause into a higher clause. If the clauses are conjoined, this should not be possible, since conjunction is a strong barrier to extraction...
[@bowern2013, 664-665]

### Examples

(@bardi1) @bowern2013, 663, example 16.64
```{r, echo=FALSE, results='asis'}
gloss_example(transliteration = "Ginyinggon roowil i-n-nya-na Ngarrigoonbooroo baali-ngan darr i-n-ar-na=jirri niimana aamba agal ambooriny Ngoolbirndi.",
             glosses = "then walk 3-TR-catch-REM.PST {NAME} shade-ALL come 3-TR-spear-REM.PST=3AUG.IO many men and people {NAME}",
             free_translation = "Then Ngarrigoonbooroo walked to her camp and came across many people at Ngoolbirndi.",
             annotation = "Ginyinggon roowil innyana Ngarrigoonbooroo baalingan darr inarnajirri niimana aamba agal ambooriny Ngoolbirndi.")
```


## Burmese
### Details
Glottocode: nucl1310 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasian

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> The sentence in (@burmese1) contains two verbs that are separated by the object of the second one, i.e. ‘hand’ [lɛʔ]. As suggested by the translation, we have here two clauses that are chained and not subordinated as in the previous example, while example (@burmese2) illustrated a SVC. Actually, in this sentence, the object of the second verb, i.e. ‘words’ [zəKa3], precedes the string of verbs, which suggests that the two verbs are closely associated syntactically. [@vittrant2012, 106]

> Vittrant (2006: 309): “Une construction de verbes en série (CVS) est une séquence de verbes qui ne sont syntaxiquement ni coordonnés ni subordonnés. Elle a d’autre part, la forme de surface d’un syntagme verbal unique.” [@vittrant2012, 106 footnote 9]

Translation: A serial verb construction is a sequence of verbes which are neither coordinated nor subordinated syntactically. It has, on the other hand, the surface form of a single verb phrase.


### Examples


(@burmese1) @vittrant2012, 107, example 5
```{r, echo=FALSE}
gloss_example(transliteration = "pəKan^3^ thɛ^3^ ye^2^ thɛ^1^ lɛʔ she^2^=Pi^3^",
             glosses = "plate inside water put hand wash=SUB.TPS",
             free_translation = "... after having poured water into the plate and having washed his hands...")
```

(@burmese2) @vittrant2012, 107, example 6
```{r, echo=FALSE}
gloss_example(transliteration = "θiN^2^bɔ^1^ pɔ^2^ =Ma^2^ θiN^2^bɔ^1^ pɔ^2^ =Ma^2^ zəKa^3^ la^2^ pyɔ^3^ =Tɛ^2^",
             glosses = "boat on LOC boat on LOC word come say REAL.ASS",
             free_translation = "On the boat, on the boat, (she) came to me to talk.",
             comment = "Lit. ...she came to tell me some words.")
```



## Bulo Stieng
### Details
Glottocode: bulo1242 <br> 
Family: Austroasiatic <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> Parmi ces critères [pour les CVS] - variables d'une langue à l'autre - les suivants ont été retenus pour être appliqués au stieng:
- le critère prosodique permettant de distinguer une CVS d'une concaténation de propositions... [@bon2014, 503]

Translation: Amongst the criteria [for SVCs] - which vary from language to language - the following have been retained for application to Stieng:
- the prosodic criterion which allows us to distinguish a SVC from a concatenation of clauses...

### Examples

> Dans l'example suivant (@bulo1), la séquence [sədiaŋ-dʔuar ɡɔk lot-kətaːŋ ʔuː rənaːp-rəsɨː] constitue un premier groupe de soufle, avec une prosodie montante. La séquence [ɓɑt joːh-ndiaŋ] constitue quant à elle un second groupe de souffle avec une intonation descendante. Ainsi, il est possible d'identifier les verbs *ɡɔk* 'ê.assis' *lot-kətaːŋ* 'ê.assis.à.genoux' comme faisant partie d'une série verbale au sein d'une même proposition; et *ɓɑt* 'plier' comme tête de la proposition suivant.
[@bon2014, 504-505]

Translation: In the following example (@bulo1), the sequence [sədiaŋ-dʔuar ɡɔk lot-kətaːŋ ʔuː rənaːp-rəsɨː] constitutes the first breath group with a rising contour. The sequence [ɓɑt joːh-ndiaŋ] in turn constitutes a second breath group with a falling intonation. Thus, it is possible to identify the verbs *ɡɔk* 'sit' *lot-kətaːŋ* 'kneel' as forming a serial verb construction within a single clause, and *ɓɑt* 'fold' as the head of the following clause.

(@bulo1) @bon2014, 505, example 839
```{r, echo=FALSE}
gloss_example(transliteration = "sədiaŋdʔuar ɡɔk lotkətaːŋ ʔuː rənaːprəsɨː (pause) ɓɑt joːhndiaŋ",
             glosses = "woman sit kneel on bamboo.mat (pause) fold shirt",
             free_translation = "The woman is kneeling on a bamboo mat and folding a shirt.")
```



## Central Khmer
### Details
Glottocode: cent1989 <br> 
Family: Austroasiatic <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

### Examples

> Citation [@haiman2011]

(@khmer1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Daakaka
### Details
Glottocode: daka1243 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: 1 <br> Grammatical: 0

### Examples

> Citation [@vonprince2012; @hopperdietzel2020]

(@daakaka1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Duhumbi
### Details
Glottocode: chug1252 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

### Examples

> Citation [@bodt2020, 437-512]

(@duhumbi1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Dzongkha
### Details
Glottocode: dzon1239 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

### Examples

> Citation [@watters2018, 138-139, 261-266]

(@dzongkha1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Eastern Arrernte
### Details
Glottocode: east2379 <br> 
Family: Pama-Nyungan <br>
Macroarea: Australian

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@wilkins1989, 508]

(@arrernte1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## English
### Details
Glottocode: stan1293 <br> 
Family: Indo-European <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

### Examples

> Citation [@huddlestonpullum2002]

(@english1) personal knowledge
```{r, echo=FALSE}
gloss_example(transliteration = "Go get me some food!",
             glosses = "go.IMP get.IMP 1SG.ACC some food")
```





## Haida
### Details
Glottocode: nort2938 <br> 
Family: Isolate <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@enrico2003]

(@haida1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Highland Populuca
### Details
Glottocode: high1276 <br> 
Family: Mixe-Zoque <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@dejongboudreault2009, 672ff.]

(@highland1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Karbi
### Details
Glottocode: karb1241 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

### Examples

> Citation [@konnerth2017, 250-258]

(@karbi1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Khroskyabs
### Details
Glottocode: guan1266 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@lai2017, 608-610]

(@khroskyabs1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Lamkang
### Details
Glottocode: lamk1238 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

### Examples

> Citation [@thounaojamchelliah2007, 65-66]

(@lamkang1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```



## Ma'anyan
### Details
Glottocode: maan1238 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@gudai1985, 284-286]

(@maanyan1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Mapudungun
### Details
Glottocode: mapu1245 <br> 
Family: Araucanian <br>
Macroarea: South America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@smeets2008, 176; @zuniga2006, 10-11, 38]

(@language1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Menya
### Details
Glottocode: meny1245 <br> 
Family: Angan <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Source [@whitehead2004, 29-30, 181-187]

(@menya1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## N|uuki
### Details
Glottocode: nuuu1241 <br> 
Family: Tuu <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

### Examples

> Citation [@collinsnamaseb2011, 53-56]

(@nuu1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Ndebele
### Details
Glottocode: nort2795 <br> 
Family: Nguni (Bantu) <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@bowernlotridge2002; @pellingpelling1974]

(@ndebele1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Nupe
### Details
Glottocode: nupe1254 <br> 
Family: Volta-Congo <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@kandybowicz2008, 54-58]

(@language1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Olo
### Details
Glottocode: oloo1241 <br> 
Family: Nuclear Torricelli <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@staley2007, 37-41]

(@olo1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Polish
### Details
Glottocode: poli1269 <br> 
Family: Indo-European <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

### Examples

> Citation [@andrason2018]

(@polish1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Sheko
### Details
Glottocode: shek1245 <br> 
Family: Dizoid <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@hellenthal2010, 334-341]

(@language1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Tahitian
### Details
Glottocode: tahi1242 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@lazardpeltzer2000, 31-33]

(@tahiti1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Tariana
### Details
Glottocode: tari1256 <br> 
Family: Arawakan <br>
Macroarea: South America

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

### Examples

> Citation [@aikhenvald1999]

(@language1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Waima'a
### Details
Glottocode: waim1251 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: 1 <br> Grammatical: 0

### Examples

> Citation Culhane p.c.

(@waimaa1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```



## Wambayan
### Details
Glottocode: wamb1258 <br> 
Family: Mirndi <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

### Examples

> Citation [@nordlinger1998, 194-195]

(@wambaya1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Western Magar
### Details
Glottocode: west218 <br> 
Family: Trans-Himalayan  <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@grunow-harsta2008, 384-387]

(@wmagar1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Yakan
### Details
Glottocode: yaka1277 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

(@yakan1) @brainardbeerens2002, 195, example 881
```{r, echo=FALSE}
gloss_example(transliteration = "pitu ku N-aka-an ka'u",
             glosses = "come ABS.1SG INTR-tell-VAL OBL.2SG",
             free_translation = "I have come to tell you.",
             annotation = "Pitu ku ngakahan ka'u.")
```





## Yongning Na
### Details
Glottocode: yong1270 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

> Most concatenations of verbs in Na fall into one of the following categories: verb phrases which belong to separate clauses but appear adjacently because their nominal arguments are not overt; a complement taking predicate with an adjacent infinitive-like verb; a verb with an aspect marker which has grammaticalized from a verb and thus has an isomorphic verb counterpart; or a resultative/cause-and-effect SVC. ... Symmetrical serial verb constructions tend to match Aktionsart marking on each verb. [@lidz2010, 396]

### Examples

(@yongningna1) @lidz2010, 397, example 566
```{r, echo=FALSE, results='asis'}
gloss_example(transliteration = "tʰi13 lə33- wæ33 zɔ33 tʰɯ33 ʐɯ33mi33 tʰɯ33 lɯ33 lə33- wɔ13 lə33- sɩ31 lə33- yɔ33 hɯ33",
             glosses = "so ACCOMP- call CSM this hearth.room this CLS ACCOMP- return ACCOMP- live ACCOMP- come FUT.REM",
             free_translation = "Having called it, the hearth room will come back to life.")
```





## Yoruba
### Details
Glottocode: yoru1245 <br> 
Family: Volta-Congo <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> ... there are a number of tests that prove the difference between cordination and the morphosyntactically more coherent SVC. One of them is extraction. While it is possible to move *ìwé* 'book' to the clausal-initial position in an SVC (@yoruba2), this is ungrammatical with coordination (@yoruba4).
[@bisang2009, 796]

### Examples

(@yoruba1) @stahlke1970, 61, example 1
```{r, echo=FALSE}
gloss_example(transliteration = "Mo mú ìwé wá ilé.",
             glosses = "1SG take book come house",
             free_translation = "I brought a book home.")
```

(@yoruba2) @bisang2009, 796, example 3'
```{r, echo=FALSE}
gloss_example(transliteration = "Ìwé ni mo mú wá ilé.",
             glosses = "book FOC 1SG take come house",
             free_translation = "It was a book I brought home.")
```

(@yoruba3) @stahlke1970, 78, example 50
```{r, echo=FALSE}
gloss_example(transliteration = "Mo mú ìwé, mo sì wá ilé.",
             glosses = "1SG take book 1SG and come home",
             free_translation = "I picked up a book and came home.")
```

(@yoruba4) @stahlke1970, 79, example 57
```{r, echo=FALSE}
gloss_example(transliteration = "Ìwé ni mo mú, mo sì wá ilé.",
             glosses = "book FOC 1SG take 1SG and come house",
             free_translation = "It was a book I took and I came home.",
             grammaticality = "*")
```

## Language
### Details
Glottocode: <br> 
Family:  <br>
Macroarea: 

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@source2001]

(@language1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```




