# Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lingglosses)
```

Here I'm going to include an overview of what data this poster is based on, including information about my coding.

## Language
### Details
Glottocode: <br> 
Family:  <br>
Macroarea: 

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@source2001]

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Afar
### Details
Glottocode: afar1241 <br> 
Family: Cushitic <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

(@afar1) @kamil2015, 291, example 907
```{r, echo=FALSE}
gloss_example(transliteration = "taamiit -íyye",
             glosses = "work.DIM say.3M.PFV",
             free_translation = "He worked a little.")
```

(@afar2) @kamil2015, 292, example 909
```{r, echo=FALSE}
gloss_example(transliteration = "boddiná mukuɖ -ítta",
             glosses = "tooth wobble say.3F.IPFV",
             free_translation = "The tooth moves.")
```

## Aimele
### Details
Glottocode: aime1238 <br> 
Family: Aimele <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Neither of the constituent verbs in the SVC in (@aimele1) is morphologically-marked as subordinate to a final verb. Instead, each verb is uninflected, except for the final verb, which bears predicate markers such as tense, aspect, mood, modality, and evidentiality. 
[@aiton2016, 201]

### Examples

(@aimele1) @aiton2016, 200, example 27
```{r, echo=FALSE}
gloss_example("koodulu suloobo -moo wɛlɛ sɛɛ -jaa",
             "downward {NAME} DAT shout say PST",
             "I shouted it down to Soloobu.")
```

(@aimele2) @aiton2016, 201, example 30
```{r, echo=FALSE}
gloss_example(transliteration = "nɛ: ti:ɸɛ: o:ga: ɛja: o:gɛ: di a:nɛ:",
              glosses = "1:SG afterwards pandanus seedling:ABS carry_in_bilum take go:PST",
              free_translation = "I went after him, taking the pandanus seeds in a bilum.")
```

(@aimele3) @aiton2016, 201, example 31
```{r, echo=FALSE}
gloss_example(transliteration = "sugu:lu: ɛna: hɛna: do:ga: dijɛ:=bi: kɛi",
               glosses = "school that:ABS go house:ABS build=IMP ASSER",
               free_translation = "Go build that building for the school!")
```

## Ami
### Details
Glottocode: ammi1238 <br> 
Family: Western Daly <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: ?

> Phonological liaison occurs between the two verbs... There are grounds here for proposing that each serial construction constitutes a single phonological conglomerate. There is no pause between the VPs and the single intonation contour points to their constituting a single clause. [@ford1998, 316]

### Examples

(@ami1) @ford1998, 316, example 8-54
```{r, echo=FALSE}
gloss_example(transliteration = "gama vulhut gama",
             glosses = "3MIN.S.R.stand beeswax 3MIN.S.R.stand",
             free_translation = "The beeswax is standing up.",
             comment = "['kamaɸ̀ʊl̪ʊkàma]")
```

(@ami2) @ford1998, 316, example 8-55
```{r, echo=FALSE}
gloss_example(transliteration = "guman-pitji gaya",
             glosses = "3MIN.A.R.poke-roll 3MIN.S.R.lie",
             free_translation = "He is still making fire.",
             comment = "['kʊmanp̀ɪcɪɣàja]")
```


## Apinayé
### Details
Glottocode: apin1244 <br> 
Family: Nuclear-Macro-Je <br>
Macroarea: South America

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: ?

> Serialization requires the nonfinite form of verbs when the (causative or instrumental morpheme) verb *ɔ* ‘do’ is involved. Otherwise, it is just the bare form of verbs that occur in a sequence; in this case, the verbs involved are usually from the class of intransitives (i.e. they take no inflection in their finite form). In Apinajé, serialization always involves either a movement verb or a position verb. Not all verbs belonging to each category are eligible for serial constructions expressing aspectual meanings, but only a selected few. When verbs from outside this inventory are used, the resulting overall meaning of the construction is more literal than grammatical, sometimes yielding awkward or jocose meanings. [@cunhadeoliveira2005, 275]

> From the prosodic point of view, the elements of each complex construction belong within a single sentential intonation contour. Morphologically, there is only one mode marker for the whole construction, such that all elements fall under its scope. In regard to syntax, boundary indices and the internally cohesive distribution of elements in the sentence are indicative of the internal coherence of serialization... [@cunhadeoliveira2005, 276]

### Examples

(@apinaye2) @cunhadeoliveira2005, 162, example 106d
```{r, echo=FALSE}
gloss_example(transliteration = "Pa tẽ tẽm.",
             glosses = "RL go fall",
             free_translation = "I fell.")
```

(@apinaye3) @cunhadeoliveira2005, 161, example 106a
```{r, echo=FALSE}
gloss_example(transliteration = "Kɔt paj arĩ ic-krĩ",
             glosses = "IRR 1.IRR stay 1-sit",
             free_translation = "I'll remain seated.")
```

(@apinaye1) @cunhadeoliveira2005, 295, example 24a
```{r, echo=FALSE}
gloss_example(transliteration = "na pa kɔt=m̃ə i-ɲ-õ pĩ katprɛ ɔ tẽ",
             glosses = "RL 1 still 1-RP-GEN wood fasten do go",
             free_translation = "I'm still walking and fastening my wood.")
```





## Ashéninka Perené
### Details
Glottocode: ashe1272 <br> 
Family: Arawakan <br>
Macroarea: South American

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

> Serialized predicates do not have more than two verbs in a verb sequence which are fully inflected grammatical and phonological words, capable of being integrated into discourse. There is no pause separating the verbs from each other... the serialized verbs are independent phonological words, each inflected for person, aspect, realis status, and mode. [@mihas2015, 163]

### Examples

(@asheninka1) @mihas2015, 165, example 6.13.b
```{r, echo=FALSE}
gloss_example(transliteration = "o- ja -t -atz -i o- shimaa -t -a",
             glosses = "3NM.S go EP PROG REAL 3NM.S fish EP REAL",
             free_translation = "She went to fish.")
```

(@asheninka2) @mihas2015, 168, example 6.19.b
```{r, echo=FALSE}
gloss_example(transliteration = "p- aako -t -ashi -t -a pi- ma -atz -i -tsi",
             glosses = "2S overtake EP APL.INT EP REAL 2S sleep PROG REAL NEG",
             free_translation = "Don't stay asleep.")
```



## Avatime!
### Details
Glottocode: avat1244 <br> 
Family: Ka-Togo <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@defina2016; @defina2016b; @vanputten2017]

(@avatime! 1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Awa-Cuaiquer
### Details
Glottocode: awac1239 <br> 
Family: Barbacoan <br>
Macroarea: South America

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: ?

> The Awa Pit Serial Verb construction thus consists of two active verbs with identical subcategorization frames, which are obligatorily adjacent and occur in a temporally iconic order. The normal range of morphological and syntactic possibilities is open to the clause, with any inflection occurring on the second verb stem. The first verb stem is either bare (if consonant-final) or suffixed with *t* (if vowel-final). Any modification, such as negation, applies semantically to both verbs, but is indicated formally on the second verb.
[@curnow1997, 305]

### Examples

(@awa1) @curnow1997, 303, example 766
```{r, echo=FALSE}
gloss_example(transliteration = "gato=na tunya pizh ku-mtu",
             glosses = "cat=TOP rat grab eat-IPFV.PART",
             free_translation = "The cat is grabbing and eating the rat.")
```

(@awa2) @curnow1997, 306, example 776
```{r, echo=FALSE}
gloss_example(transliteration = "pa azh kway-zi",
             glosses = "sun shine drop-NONLOCUT",
             free_translation = "The sun shone out.")
```

(@awa3) @curnow1997, 306, example 775
```{r, echo=FALSE}
gloss_example(transliteration = "Demetrio=na I-t kway-zi",
             glosses = "Demetrio=TOP go-SV drop-NONLOCUT",
             free_translation = "Demetrio left.")
```

## Baure
### Details
Glottocode: baur1253 <br> 
Family: Arawakan <br>
Macroarea: South America

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: ?

> Finite verbs with possible person cross-reference on the verb, as in Baure, may also occur in serial verb constructions. ... The coordination of predicates is identical with the coordination of clauses without connector. Serial verb constructions, on the other hand, are a number of verbs that are uttered in one clause. In Baure the two categories seem pretty well distinguishable... [@danielsen2007, 427]

### Examples

(@baure1) @danielsen2007, 428, example 144
```{r, echo=FALSE}
gloss_example(transliteration = "ro=kač-po-wo ro=pino-po",
             glosses = "3SG.M=go-PRFLX-COP 3SG.M=flee-PRFLX",
             annotation = "*rokačpow* *ropinop*",
             free_translation = "he (went away and) fled")
```

## Bardi
### Details
Glottocode: bard1255 <br> 
Family: Nyulnyulan <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> [Serial verbs] occur in a single intonation contour (although there are also examples with breaks, and examples where intonation units and syntactic units are not isomorphic). 
[@bowern2013, 663-664]

> A further piece of evidence shows that these clauses are joined in a single higher structure rather than chained without embedding. It is possible to extract constituents from a lower clause into a higher clause. If the clauses are conjoined, this should not be possible, since conjunction is a strong barrier to extraction...
[@bowern2013, 664-665]

### Examples

(@bardi1) @bowern2013, 663, example 16.64
```{r, echo=FALSE, results='asis'}
gloss_example(transliteration = "Ginyinggon roowil i-n-nya-na Ngarrigoonbooroo baali-ngan darr i-n-ar-na=jirri niimana aamba agal ambooriny Ngoolbirndi.",
             glosses = "then walk 3-TR-catch-REM.PST {NAME} shade-ALL come 3-TR-spear-REM.PST=3AUG.IO many men and people {NAME}",
             free_translation = "Then Ngarrigoonbooroo walked to her camp and came across many people at Ngoolbirndi.",
             annotation = "Ginyinggon roowil innyana Ngarrigoonbooroo baalingan darr inarnajirri niimana aamba agal ambooriny Ngoolbirndi.")
```

## Burmese
### Details
Glottocode: nucl1310 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasian

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> The sentence in (@burmese1) contains two verbs that are separated by the object of the second one, i.e. ‘hand’ [lɛʔ]. As suggested by the translation, we have here two clauses that are chained and not subordinated as in the previous example, while example (@burmese2) illustrated a SVC. Actually, in this sentence, the object of the second verb, i.e. ‘words’ [zəKa3], precedes the string of verbs, which suggests that the two verbs are closely associated syntactically. [@vittrant2012, 106]

<br>

> Vittrant (2006: 309): “Une construction de verbes en série (CVS) est une séquence de verbes qui ne sont syntaxiquement ni coordonnés ni subordonnés. Elle a d’autre part, la forme de surface d’un syntagme verbal unique.” [@vittrant2012, 106 footnote 9]

Translation: A serial verb construction is a sequence of verbes which are neither coordinated nor subordinated syntactically. It has, on the other hand, the surface form of a single verb phrase.

### Examples

(@burmese1) @vittrant2012, 107, example 5
```{r, echo=FALSE}
gloss_example(transliteration = "pəKan^3^ thɛ^3^ ye^2^ thɛ^1^ lɛʔ she^2^=Pi^3^",
             glosses = "plate inside water put hand wash=SUB.TPS",
             free_translation = "... after having poured water into the plate and having washed his hands...")
```

(@burmese2) @vittrant2012, 107, example 6
```{r, echo=FALSE}
gloss_example(transliteration = "θiN^2^bɔ^1^ pɔ^2^ =Ma^2^ θiN^2^bɔ^1^ pɔ^2^ =Ma^2^ zəKa^3^ la^2^ pyɔ^3^ =Tɛ^2^",
             glosses = "boat on LOC boat on LOC word come say REAL.ASS",
             free_translation = "On the boat, on the boat, (she) came to me to talk.",
             comment = "Lit. ...she came to tell me some words.")
```



## Bulo Stieng
### Details
Glottocode: bulo1242 <br> 
Family: Austroasiatic <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> Parmi ces critères [pour les CVS] - variables d'une langue à l'autre - les suivants ont été retenus pour être appliqués au stieng:
- le critère prosodique permettant de distinguer une CVS d'une concaténation de propositions... [@bon2014, 503]

Translation: Amongst the criteria [for SVCs] - which vary from language to language - the following have been retained for application to Stieng:
- the prosodic criterion which allows us to distinguish a SVC from a concatenation of clauses...

### Examples

> Dans l'example suivant (@bulo1), la séquence [sədiaŋ-dʔuar ɡɔk lot-kətaːŋ ʔuː rənaːp-rəsɨː] constitue un premier groupe de soufle, avec une prosodie montante. La séquence [ɓɑt joːh-ndiaŋ] constitue quant à elle un second groupe de souffle avec une intonation descendante. Ainsi, il est possible d'identifier les verbs *ɡɔk* 'ê.assis' *lot-kətaːŋ* 'ê.assis.à.genoux' comme faisant partie d'une série verbale au sein d'une même proposition; et *ɓɑt* 'plier' comme tête de la proposition suivant.
[@bon2014, 504-505]

Translation: In the following example (@bulo1), the sequence [sədiaŋ-dʔuar ɡɔk lot-kətaːŋ ʔuː rənaːp-rəsɨː] constitutes the first breath group with a rising contour. The sequence [ɓɑt joːh-ndiaŋ] in turn constitutes a second breath group with a falling intonation. Thus, it is possible to identify the verbs *ɡɔk* 'sit' *lot-kətaːŋ* 'kneel' as forming a serial verb construction within a single clause, and *ɓɑt* 'fold' as the head of the following clause.

(@bulo1) @bon2014, 505, example 839
```{r, echo=FALSE}
gloss_example(transliteration = "sədiaŋdʔuar ɡɔk lotkətaːŋ ʔuː rənaːprəsɨː (pause) ɓɑt joːhndiaŋ",
             glosses = "woman sit kneel on bamboo.mat (pause) fold shirt",
             free_translation = "The woman is kneeling on a bamboo mat and folding a shirt.")
```



## Burarra
### Details
Glottocode: bura1267 <br> 
Family: Maningrida <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

Note that glosses have been adapted for a general audience.

(@burarra2) @green1987, 79, example 5(lxxxii)
```{r, echo=FALSE}
gloss_example(transliteration = "Nginyi-pa gornabola ø-gana-ø ø-bamba-ø",
             glosses = "2MIN-DIR wallaby 2MIN.IMP-watch-IRR 2MIN.IMP-go_along-IRR",
             free_translation = "You go along watching for wallabies.")
```

(@burarra3) @green1987, 80, example 5(lxxxix)
```{r, echo=FALSE}
gloss_example(transliteration = "ngu-durtchi-nga ngu-bo-na",
             glosses = "1MIN-be_full-RL 1MIN-go-ANT",
             free_translation = "I'm full.")
```

(@burarra1) Green 1987:77, example 5(lxxviii)
```{r, echo=FALSE}
gloss_example(transliteration = "m-ba-nga a-rrik-nga gorrngunya",
             glosses = "3MIN>3MIN.mun-eat-PST 3MIN.an-crawl-PST grass",
             free_translation = "He crawled along eating grass.")
```

## Central Khmer!
### Details
Glottocode: cent1989 <br> 
Family: Austroasiatic <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

### Examples

> Citation [@haiman2011]

(@khmer1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Daakaka!
### Details
Glottocode: daka1243 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: 1 <br> Grammatical: 0

### Examples

> Citation [@vonprince2012; @hopperdietzel2020]

(@daakaka1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Duhumbi!
### Details
Glottocode: chug1252 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

### Examples

> Citation [@bodt2020, 437-512]

(@duhumbi1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Dzongkha!
### Details
Glottocode: dzon1239 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

### Examples

> Citation [@watters2018, 138-139, 261-266]

(@dzongkha1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Eastern Arrernte!
### Details
Glottocode: east2379 <br> 
Family: Pama-Nyungan <br>
Macroarea: Australian

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@wilkins1989, 508]

(@arrernte1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## English
### Details
Glottocode: stan1293 <br> 
Family: Indo-European <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

### Examples

> Citation [@huddlestonpullum2002]

(@english1) personal knowledge
```{r, echo=FALSE}
gloss_example(transliteration = "Go get me some food!",
             glosses = "go.IMP get.IMP 1SG.ACC some food")
```





## Haida!
### Details
Glottocode: nort2938 <br> 
Family: Isolate <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@enrico2003]

(@haida1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Hakhun Tangsa!
### Details
Glottocode: noct1238 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

There is a clear morphosyntactic template for verbs. 

### Examples

> Citation [@boro2017]

(@hakhun1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Highland Populuca!
### Details
Glottocode: high1276 <br> 
Family: Mixe-Zoque <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@dejongboudreault2009, 672ff.]

(@highland1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Karbi!
### Details
Glottocode: karb1241 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

### Examples

> Citation [@konnerth2017, 250-258]

(@karbi1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Kayardild
### Details
Glottocode: kaya1319 <br> 
Family: Tangkic <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

> In (@kayardild2) the verbs belong to different clauses, as shown by the independent intonation contours (indicated here by a comma); a literal meaning is forced. ... In (@kayardild3) ... *thaatha* is a clausemate and follows the main verb, giving the 'go and return' meaning. [@evans1995, 308]

### Examples

(@kayardild1) @evans1995, 306-307, example 8-40
```{r, echo=FALSE}
gloss_example(transliteration = "nyingka dathin-ki dulk-i ngadi-ja kuluu-j, nyingka kalka-th",
             glosses = "2SG.NOM that-MLOC place-MLOC wound-ACT scratch-ACT 2SG.NOM sicken-ACT",
             free_translation = "You have despoiled that place by digging there, and you have fallen ill.")
```

(@kayardild2) @evans1995, 308, example 8-45
```{r, echo=FALSE}
gloss_example(transliteration = "niya kuujuu-j, thaa-th",
             glosses = "3SG.NOM swim-ACT return-ACT",
             free_translation = "He had a swim, and came back.")
```

(@kayardild3) @evans1995, 308, example 8-47
```{r, echo=FALSE}
gloss_example(transliteration = "niya kuujuu-ja thaa-th",
             glosses = "3SG.NOM swim-ACT return-ACT",
             free_translation = "He's gone for a swim (and will come back).")
```

## Khroskyabs!
### Details
Glottocode: guan1266 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@lai2017, 608-610]

(@khroskyabs1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Lamkang!
### Details
Glottocode: lamk1238 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

### Examples

> Citation [@thounaojamchelliah2007, 65-66]

(@lamkang1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```



## Ma'anyan!
### Details
Glottocode: maan1238 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@gudai1985, 284-286]

(@maanyan1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Mapudungun!
### Details
Glottocode: mapu1245 <br> 
Family: Araucanian <br>
Macroarea: South America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@smeets2008, 176; @zuniga2006, 10-11, 38]

(@language1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Mawng
### Details
Glottocode: maun1240 <br> 
Family: Iwaidjan <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

Although there is some evidence for the serial verb construction forming a grammatical unit [@singer2006, 132], there is no grammatical marking involved which could potentially signal the continuation of a syntactic unit across a pause.

### Examples

(@mawng1) @singer2006, 131, example 4-85
```{r, echo=FALSE, results='asis'}
gloss_example(transliteration = "Waka-pa k-i-wani-ø k-ani-mirlaju-n ja y-ara ja single.",
             glosses = "DIST.LAND-EMPH PRS-3MASC-sit-NPST PRS-3MASC/3LAND-wait_for-NPST MASC MASC-other MASC single_man",
             free_translation = "During that time he keeps on waiting (for a wife), the single man.")
```

(@mawng2) @singer2006, 131, example 4-86
```{r, echo=FALSE}
gloss_example(transliteration = "K-a-wani-ø apu-la-ngung walij",
             glosses = "PRS-3PL-sit-NPST 3PL/3EDIB-eat-PST.CONT food",
             free_translation = "They were eating food.")
```

(@mawng3) @singer2006, 132, example 4-87
```{r, echo=FALSE}
gloss_example(transliteration = "Nuka ja k-i-wani-ø k-ani-wurru-n.",
             glosses = "PROX.MASC MASC PRS-3MASC-sit-NPST PRS-3MASC/3LAND-know-NPST",
             free_translation = "He's the one who knows.")
```

## Menya!
### Details
Glottocode: meny1245 <br> 
Family: Angan <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Source [@whitehead2004, 29-30, 181-187]

(@menya1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Murrinhpatha
### Details
Glottocode: murr1258 <br> 
Family: Southern Daly <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: ?

### Examples

(@murrinhpatha2) @blythe2009, 364, On the Flat line 103
```{r, echo=FALSE}
gloss_example(transliteration = "ngunu-murrk-nu=ngi",
             glosses = "1SG.S.7go.FUT-eat-FUT=1SG.S.1sit.FUT",
             free_translation = "I want to eat some.")
```

(@murrinhpatha3) @blythe2009, 378, Ningbingi line 13
```{r, echo=FALSE}
gloss_example(transliteration = "me-nge-dha=pirri",
             glosses = "3SG.S.say.PSTIMP-3SG.IO-PST=3SG.S.3stand.PSTIMP",
             free_translation = "He was saying to her...")
```

(@murrinhpatha1) @blythe2009, 364, On the Flat line 87
```{r, echo=FALSE, results = 'asis'}
gloss_example(transliteration = "ku panda thanggugu thurdana-ngintha-yitj-tha=tjini ha ha",
             glosses = "animate that.you.know what_animate? 2SG.S.30.PSTIMP-DU.F.NSIB-tell_story-PST=2SG.S.1sit.PSTIMP ha ha",
             free_translation = "That thing of the 'ku' class, whatitsname. You two were telling that story. ")
```


## N|uuki!
### Details
Glottocode: nuuu1241 <br> 
Family: Tuu <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

### Examples

> Citation [@collinsnamaseb2011, 53-56]

(@nuu1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Ndebele!
### Details
Glottocode: nort2795 <br> 
Family: Nguni (Bantu) <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@bowernlotridge2002; @pellingpelling1974]

(@ndebele1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Nupe!
### Details
Glottocode: nupe1254 <br> 
Family: Volta-Congo <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@kandybowicz2008, 54-58]

(@language1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Olo!
### Details
Glottocode: oloo1241 <br> 
Family: Nuclear Torricelli <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@staley2007, 37-41]

(@olo1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Polish!
### Details
Glottocode: poli1269 <br> 
Family: Indo-European <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

### Examples

> Citation [@andrason2018]

(@polish1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Sheko!
### Details
Glottocode: shek1245 <br> 
Family: Dizoid <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@hellenthal2010, 334-341]

(@language1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Tahitian!
### Details
Glottocode: tahi1242 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@lazardpeltzer2000, 31-33]

(@tahiti1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Tariana!
### Details
Glottocode: tari1256 <br> 
Family: Arawakan <br>
Macroarea: South America

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

### Examples

> Citation [@aikhenvald1999]

(@language1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Waima'a!
### Details
Glottocode: waim1251 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: 1 <br> Grammatical: 0

### Examples

> Citation Culhane p.c.

(@waimaa1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```



## Wambayan!
### Details
Glottocode: wamb1258 <br> 
Family: Mirndi <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

### Examples

> Citation [@nordlinger1998, 194-195]

(@wambaya1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Western Magar!
### Details
Glottocode: west218 <br> 
Family: Trans-Himalayan  <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@grunow-harsta2008, 384-387]

(@wmagar1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Yakan
### Details
Glottocode: yaka1277 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

(@yakan1) @brainardbeerens2002, 195, example 881
```{r, echo=FALSE}
gloss_example(transliteration = "pitu ku N-aka-an ka'u",
             glosses = "come ABS.1SG INTR-tell-VAL OBL.2SG",
             free_translation = "I have come to tell you.",
             annotation = "Pitu ku ngakahan ka'u.")
```





## Yongning Na
### Details
Glottocode: yong1270 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

> Most concatenations of verbs in Na fall into one of the following categories: verb phrases which belong to separate clauses but appear adjacently because their nominal arguments are not overt; a complement taking predicate with an adjacent infinitive-like verb; a verb with an aspect marker which has grammaticalized from a verb and thus has an isomorphic verb counterpart; or a resultative/cause-and-effect SVC. ... Symmetrical serial verb constructions tend to match Aktionsart marking on each verb. [@lidz2010, 396]

### Examples

(@yongningna1) @lidz2010, 397, example 566
```{r, echo=FALSE, results='asis'}
gloss_example(transliteration = "tʰi13 lə33- wæ33 zɔ33 tʰɯ33 ʐɯ33mi33 tʰɯ33 lɯ33 lə33- wɔ13 lə33- sɩ31 lə33- yɔ33 hɯ33",
             glosses = "so ACCOMP- call CSM this hearth.room this CLS ACCOMP- return ACCOMP- live ACCOMP- come FUT.REM",
             free_translation = "Having called it, the hearth room will come back to life.")
```





## Yoruba
### Details
Glottocode: yoru1245 <br> 
Family: Volta-Congo <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> ... there are a number of tests that prove the difference between cordination and the morphosyntactically more coherent SVC. One of them is extraction. While it is possible to move *ìwé* 'book' to the clausal-initial position in an SVC (@yoruba2), this is ungrammatical with coordination (@yoruba4).
[@bisang2009, 796]

### Examples

(@yoruba1) @stahlke1970, 61, example 1
```{r, echo=FALSE}
gloss_example(transliteration = "Mo mú ìwé wá ilé.",
             glosses = "1SG take book come house",
             free_translation = "I brought a book home.")
```

(@yoruba2) @bisang2009, 796, example 3'
```{r, echo=FALSE}
gloss_example(transliteration = "Ìwé ni mo mú wá ilé.",
             glosses = "book FOC 1SG take come house",
             free_translation = "It was a book I brought home.")
```

(@yoruba3) @stahlke1970, 78, example 50
```{r, echo=FALSE}
gloss_example(transliteration = "Mo mú ìwé, mo sì wá ilé.",
             glosses = "1SG take book 1SG and come home",
             free_translation = "I picked up a book and came home.")
```

(@yoruba4) @stahlke1970, 79, example 57
```{r, echo=FALSE}
gloss_example(transliteration = "Ìwé ni mo mú, mo sì wá ilé.",
             glosses = "book FOC 1SG take 1SG and come house",
             free_translation = "It was a book I took and I came home.",
             grammaticality = "*")
```
