```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lingglosses)
#library(linguisticsdown)
#linguisticsdown::writeIPA()
```

## Haida!
### Details
Glottocode: nort2938 <br> 
Family: Isolate <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@enrico2003]

(@haida1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Hakhun Tangsa!
### Details
Glottocode: noct1238 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

There is a clear morphosyntactic template for verbs. 

### Examples

> Citation [@boro2017]

(@hakhun1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Hano
### Details
Glottocode: hano1246 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Haruai
### Details
Glottocode: haru1245 <br> 
Family: Piawi <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Hatam
### Details
Glottocode: hata1243 <br> 
Family: Hatam-Mansim <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Havasupai-Walapai-Yavapai
### Details
Glottocode: hava1248 <br> 
Family: Cochimi-Yuman <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Hidatsa
### Details
Glottocode: hida1246 <br> 
Family: Siouan  <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> There are two types of serial verb constructions in Hidatsa. In the first type (the majority of cases), except for the final verb in the sequence, all preceding component-verbs are marked with the contemporaneous suffix that triggers ablaut on the immediately preceding vowel. In the other type, the nonfinal verbs are not marked with the contemporaneous suffix and so do not ablaut. ... Subjects of verbs in serial verb constructions are coreferential and, with the exception of component-verbs that precede motion verbs in unmarked serial constructions, usually marked with pronominal prefixes. In the normal rate of speech serial verb constructions *usually* [emphasis NP] form a pitch phrase. [@park2012, 538-539]

> Contemporaneous serial verb constructions are always paraphrasable into two clauses that are then conceptualized as two separate events; i.e., replacing the contemporaneous suffix with the coordinative suffix yields two clauses that are conceptualized as two separate events (typically, events in sequence). [@park2012, 541]

> None of the elements in an unmarked serial construction are marked with ablaut-triggering contemporaneous suffixes if the semantics of the serialized construction encodes intent, desire, or purpose. The so-called purposive serial verb constructions anticipate future events, whereby the intended action follows the intent or motion encoded in the second verb. [@park2012, 544]

Note that the final verb in a clause *must* take an illocutionary suffix, which is non-zero except for imperatives to single subjects [@park2012, 224-225].

### Examples

(@hidatsa1) @park2012, 540, example 111
```{r, echo=FALSE}
#gloss_example(transliteration = "miidí-Ø nuwí-Ø úˀsia-c",
#             glosses = "crawl-CONT go.around-CONT arrive-DECL",
#             free_translation = "He arrived by crawling.",
#             annotation = "Miidá nuwá úˀsiaˀc.")
```
<!-- fix example -->


(@hidatsa2) @park2012, 545, example 124
```{r, echo=FALSE}
#gloss_example(transliteration = "ma-iréˀ goowí-waa-c",
#             glosses = "1POS-speak finish-1CAUS.DIR-DECL",
#             free_translation = "I finished talking.",
#             annotation = "Maréˀ goowíwaac.")
```
<!-- fix example -->

## Highland Populuca!
### Details
Glottocode: high1276 <br> 
Family: Mixe-Zoque <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@dejongboudreault2009, 672ff.]

(@highland1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "a",
             glosses = "word",
             free_translation = "here")
```





## Hinuq
### Details
Glottocode: hinu1240 <br> 
Family: Nakh-Dagestanian <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> Serial verb constructions in Hinuq consist of two juxtaposed verbs which act together as a single predicate. The first verb appears only as a stem without any further suffixes but retains its agreement prefixes. The second verb takes all inflectional and derivational suffixes that apply to the whole complex. ... Many serial verbs belong to the semantic domain of verbs of motion and physical actions, but they cannot be characterized by other semantic criteria (e.g. they do not follow iconic principles, they do not denote culturally important concatenations of events). Serial verbs cannot be regarded as morphosyntactic devices (except for those expressing causative constructions). [@forker2013, 697]

> Nakh-Daghestanian languages do not belong to the languages that are generally known to employ verb serialization. Nevertheless, Hinuq serial verb constructions match well with Aikhenvald’s (2006) definition. They always consist of two verbs that together form a single predicate and have one single intonation contour. The two verbs are phonologically independent words, but they form one grammatical word, as can be seen by the argument structure and the agreement. Serial verbs share all arguments and modifiers and the agreement. [@forker2013, 702]

### Examples

(@hinuq1) @forker2013, 699, example 1283a
```{r, echo=FALSE}
#gloss_example(transliteration = "hayɬoy ɣʷero-z t’oq hat͡ɬi-cat͡ɬi-š",
#             glosses = "he.ERG cow-DAT knife bump-throw-PST",
#             free_translation = "He repeatedly hit the cow with the knife.")
```
<!-- fix example -->

(@hinuq1) @forker2013, 700, example 1285b
```{r, echo=FALSE}
#gloss_example(transliteration = "hibayɬi kek-Ø-ece goɬi polnostyu kek-Ø-ece goɬi de",
#             glosses = "there break-I-melt be.IRR completely break-I-melt be.IRR 1SG",
#             free_translation = "There I (masc.) would be smashed off, completely smashed off.")
```
<!-- fix example -->

## Ho
### Details
Glottocode: hooo1248 <br> 
Family: Austroasiatic <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

> We will see that Ho has nuclear serialization. The verbs in a nuclear serialization form a tight bond; there is only one set of arguments, and aspect marking modifies the complex event (rather than each verb root separately). ... We see that serialization is a productive feature of Ho grammar and, additionally, that many common V2s have grammaticalized to varying degrees. [@pucilowski2013, 202]

> In a serial verb construction, both verb roots form part of the same predicate i.e., there is no subordinating suffix, and they share subject clitics and object suffixes. Importantly, the negative particle and subject clitic – both of which must appear immediately before the main verb precede the entire serial verb construction, illustrating that it is a single unit. [@pucilowski2013, 203]

### Examples

(@ho1) @pucilowski2013, 204, example 6.55
```{r, echo=FALSE}
#gloss_example(transliteration = "aɲ gaɽa-te iɖi-atu:-ɲ-pe",
#             glosses = "1SG river-ALL take-leave-1SG-IMP.2PL",
#             free_translation = "take me to the river and leave me")
```
<!-- fix examples -->

(@ho1) @pucilowski2013, 204, example 6.55
```{r, echo=FALSE}
#gloss_example(transliteration = "naʔ=do miyaɖ baru_daru baʈi-gur-ke-ɖ-a, maʔ-gur-ke-ɖ-a=eʔ",
#             glosses = "now=FOC one.INAN kusum_tree overturn-fall-PFV-TR-FIN cut_swing-fall-PFV-TR-FIN=3SG",
#             free_translation = "now he brought the kusum tree down, [he] chopped it down")
```
<!-- fix examples -->

## Huli
### Details
Glottocode: huli1244 <br> 
Family: Nuclear Trans-New-Guinea <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Hup
### Details
Glottocode: hupd1244 <br> 
Family: Naduhup <br>
Macroarea: South America

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> The Hup verb compound is here understood as a type of serial verb construction that involves contiguous ‘nuclear’ serialization, resulting in a single phonological word ... In those Hup compounds that follow clear iconic principles, verb order corresponds to temporal succession of actions, including sub-events in a cause-effect relationship. Those Hup compounds that are less iconic (though perhaps not completely non-iconic, as discussed below) encode coincident motion, posture, and manner ... Finally, Hup verb compounds do not allow duplicate participant roles, such as two agents or two objects...
[@epps2008, 393]

> As a single-predicate, cosubordinate construction, a compound verb in Hup must encode an event that has a degree of conceptual unity. When this unity is not judged to be present, the events are necessarily represented by two predicates, linked by subordination or coordination strategies. However, among compound verbs, the linked stems can encode a variety of more or less conceptually integrated sub-events. [@epps2008, 395]

### Examples

(@hup1) @epps2008, 396, example 9
```{r, echo=FALSE}
#gloss_example(transliteration = "ʔæ̌ytæn=yɨ́ʔ hɨd mæh-b’uy-d’əh-ham-yɨ́ʔ-ay-áh!",
#             glosses = "together=TEL 3PL kill-throw-send-go-TEL-INCH-DECL",
#             free_translation = "Together they killed them and threw them out!")
```
<!-- fix examples -->

(@hup2) @epps2008, 397, example 15
```{r, echo=FALSE}
#gloss_example(transliteration = "tɨ́h-ǎn kayǎk tɔ́ʔ hɨd g’ɔʔ-nɔ́ʔ-ɔ̃́y",
#             glosses = "3SG-OBJ manioc tuber 3PL pull.up-give-DYNM",
#             free_translation = "They’ve pulled up manioc and given it to her.")
```

Please note that not all diacritics could be represented accurately in (@hup3); please consult original source for original example.

(@hup3) @epps2008, 400, example 25
```{r, echo=FALSE}
#gloss_example(transliteration = "tɨh yok-d’əh-ham-yæh-kamí=mah",
#             glosses = "3SG poke-send-go-FRUST-time_of=REP",
#             free_translation = "At the moment when she poked (the knife) in...")
```
<!-- fix examples -->

## Iatmul
### Details
Glottocode: iatm1242 <br> 
Family: Ndu <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Imonda
### Details
Glottocode: imon1245 <br> 
Family: Border <br>
Macroarea: Papua New Guinea

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Irabu
### Details
Glottocode: irab1238 <br> 
Family: Japonic <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> There are two major differences between compound verbs and agglutinative SVCs in Irabu. First, V2 in an SVC never undergoes sequential voicing, while V2 in a compound may do so ... Second, unlike compounds, V1 and V2 in an SVC constitute separate primary stems. Since each stem is a primary stem, each can carry its own derivational affix as long as it is semantically appropriate. As a single word, however, the serialised stems carry a single inflectional affix. [@shimoji2008, 302-303]

### Examples

(@irabu1) @shimoji2008, 303, example 6-71
```{r, echo=FALSE}
#gloss_example(transliteration = "ibi-pazїmi-sїmi-tar",
#             glosses = "plant-begin-CAUS-PST",
#             free_translation = "ordered someone to begin planting")
```


## I'saka
### Details
Glottocode: kris1246 <br> 
Family: Sko <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Jamaican Creole English
### Details
Glottocode: jama1262 <br> 
Family: Indo-European <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> SVCs involve two or more verbs brought together without a complementizer, conjunction or infinitive marker, and with no pause. If TMA or negation are marked, the marking on all verbs agrees, and typically only occurs on the first. There is normally a single expressed subject, and one direct object (if any)... [@patrick2008]

> In Jamaican SVCs, it is also possible to focus the verbs in an SVC separately through predicate clefting ... [@hollington2015, 164]

### Examples

(@jcreole1) @hollington2015, 162, example 214
```{r, echo=FALSE}
gloss_example(transliteration = "Anansi iit kyari uom di pat",
             glosses = "Anansi eat carry home DEF pot",
             free_translation = "Anansi ate and carried the pot home.")
```

(@jcreole2) @hollington2015, 164, example 223
```{r, echo=FALSE}
gloss_example(transliteration = "A gi im bai di buk gi im",
             glosses = "COP give 3SG buy DEF book give 3SG",
             free_translation = "(S)he bought the book *for* him/her.")
```

(@jcreole3) @hollington2015, 164, example 223
```{r, echo=FALSE}
gloss_example(transliteration = "A bai im bai di buk gi im",
             glosses = "COP buy 3SG buy DEF book give 3SG",
             free_translation = "(S)he *bought* the book for him/her.")
```
