```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lingglosses)
#library(linguisticsdown)
#linguisticsdown::writeIPA()
```

## Bafut
### Details
Glottocode: bafu1246 <br> 
Family: Grassfields <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> In the serial verb construction in Bafut, two or more verbs which all refer to subparts or aspects of a single overall event occur in a series in the same construction. The second verb usually denotes an outgrowth of the action denoted by the first verb and represents a further development, consequence, result, goal or culmination of the action of the first verb. Morphologically, the serial verb construction (henceforth SVC) is marked by an initial homorganic nasal consonant on the second verb (or subsequent verbs) when the construction is in the past tense. [@tamanji2010, 182]

### Examples

(@bafut1) @tamanji2010, 183, example 28c
```{r, echo=FALSE, results='asis'}
#gloss_example(transliteration = "n-ó j-á=ɨ̀ kɨ̀ kóŋnə̂ N-ʒÎ N-lóô ǹ-dɨ̀ɨ̀ w-á",
#             glosses = "9-snake 9-the=SM P2 silently N-come N-bite 1-witch 1-the",
#             free_translation = "The snake silently came up and bit the witch.",
#             annotation = "nó já kɨ̀ kòŋnɨ̀ nʒì nlóó ndɨ̀ɨ̀ wá.")
```

(@bafut2) @tamanji2010, 183, example 29b
```{r, echo=FALSE}
#gloss_example(transliteration = "ǹ-tsìì-twúgɨ̀ w-á-à kɨ̀ bwíì n-lì'ìnə́ n-óò",
#             glosses = "1-guard-night 1-the=SM P2 sleep N-forget 1-time",
#             free_translation = "The snake silently came up and bit the witch.",
#             annotation = "ǹtsììtwúgɨ̀ wá kɨ̀ bwíì nlì'ìnə́ nóò.")
```

## Bambassi
### Details
Glottocode: bamb1262 <br> 
Family: Blue Nile Mao <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

> Northern Mao natural discourse exhibits only a very occasional use of serialization, i.e. a construction in which verb stems occur in immediate succession and where the subjects of each of these verbs is the same, but where none of the verbs is marked with the SS:NF suffix or any other dependency marker. In contrast, verbal compounds, where two verb stems (i.e. roots with verbal tone melodies) join to form a single complex verb stem, the entirety of which is surrounded by inflectional prefixes and suffixes, are more commonly attested. These compounds form single morpho-phonological words. Verbal compounding mainly involves a grammaticalized verb from a small closed set of verbs joining with a lexical verb from a large open class. [@ahland2012, 601]

### Examples

(@bambassi1) @ahland2012, 602, example 12.73
```{r, echo=FALSE}
gloss_example(transliteration = "kí mí-in tí-hów-j-à",
             glosses = "come eat-SS.NF 1SG-go-AWAY-DECL",
             free_translation = "I came, ate, and left.")
```

(@bambassi2) @ahland2012, 603, example 12.78
```{r, echo=FALSE}
gloss_example(transliteration = "pàpáj kez-èt pór k'íl-kòt'-in tí-hów-kj-á",
             glosses = "papaya top-LOC forget leave-PF-SS.NF 1SG-go-TOWARD-DECL",
             free_translation = "I forgot and have left (it) up in the papaya tree and came.")
```

## Bangime
### Details
Glottocode: bang1363 <br> 
Family: Isolate <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Barai
### Details
Glottocode: nucl1630 <br> 
Family: Koiarian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: 1 <br> Grammatical: 0

> Another important difference between [tight and loose] serial verb constructions has to do with intonation. We noted earlier that both grammatical and phonological features coincide to mark the boundaries between sentences in Barai (and Papuan languages generally) and that the significant phonological feature was an overall intonation pattern superimposed over the entire clause chain. ... The relevant fact for our consideration here is that the standard serial construction requires the phrasal contour over each conjunct of its series. The more tightly knit string of juxtaposed predicates collectively takes just a single phrasal contour. [@olson1981, 176-177]

### Examples

(@barai1) @olson1981, 177, example 57
```{r, echo=FALSE}
gloss_example(transliteration = "na fi fase isoe-vo",
             glosses = "1SG sit letter write-PRS",
             free_translation = "I am sitting and writing a letter.")
```

(@barai2) @olson1981, 177, example 58
```{r, echo=FALSE}
gloss_example(transliteration = "na fase fi isoe-vo",
             glosses = "1SG letter sit write-PRS",
             free_translation = "I am sitting writing a letter.")
```

## Bardi
### Details
Glottocode: bard1255 <br> 
Family: Nyulnyulan <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> [Serial verbs] occur in a single intonation contour (although there are also examples with breaks, and examples where intonation units and syntactic units are not isomorphic). 
[@bowern2013, 663-664]

> A further piece of evidence shows that these clauses are joined in a single higher structure rather than chained without embedding. It is possible to extract constituents from a lower clause into a higher clause. If the clauses are conjoined, this should not be possible, since conjunction is a strong barrier to extraction...
[@bowern2013, 664-665]

### Examples

(@bardi1) @bowern2013, 663, example 16.64
```{r, echo=FALSE, results='asis'}
gloss_example(transliteration = "Ginyinggon roowil i-n-nya-na Ngarrigoonbooroo baali-ngan darr i-n-ar-na=jirri niimana aamba agal ambooriny Ngoolbirndi.",
             glosses = "then walk 3-TR-catch-REM.PST {NAME} shade-ALL come 3-TR-spear-REM.PST=3AUG.IO many men and people {NAME}",
             free_translation = "Then Ngarrigoonbooroo walked to her camp and came across many people at Ngoolbirndi.",
             annotation = "Ginyinggon roowil innyana Ngarrigoonbooroo baalingan darr inarnajirri niimana aamba agal ambooriny Ngoolbirndi.")
```

## Batak Karo
### Details
Glottocode: bata1293 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> A favourite clause-combining strategy in Karo is serialisation, whereby two clauses are welded into one by virtue of a shared nominal element (normally the subject). [@woollams1996, 325]

> It is very common in Karo for two independent clauses to be combined in a single sentence without the use of an explicit connective. This conjoining into the one sentence is signalled phonologically by a shorter than usual intersentential pause, and orthographically by a comma instead of a full stop. [@woollams1996, 344]

### Examples

(@karo1) @woollams1996, 327, example 8.213
```{r, echo=FALSE}
gloss_example(transliteration = "Naktak iluh Datuk Rubia Gandé megi sora kempu-na rendé.",
             glosses = "fall tears Datuk Rubia Gande hear voice grandson-3.POSS sing",
             free_translation = "Datuk Rubia Gande shed tears as he heard his grandson singing.")
```

(@karo1) @woollams1996, 328, example 8.219
```{r, echo=FALSE}
gloss_example(transliteration = "Asuh-asuh-en-na enggo keri ben-én i-tangko kalak.",
             glosses = "tend-RDP-PL-3.POSS already gone lost-PL PASS-steal person",
             free_translation = "Their cattle were all gone, missing, stolen by somebody.")
```

## Baure
### Details
Glottocode: baur1253 <br> 
Family: Arawakan <br>
Macroarea: South America

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: ?

> Finite verbs with possible person cross-reference on the verb, as in Baure, may also occur in serial verb constructions. ... The coordination of predicates is identical with the coordination of clauses without connector. Serial verb constructions, on the other hand, are a number of verbs that are uttered in one clause. In Baure the two categories seem pretty well distinguishable... [@danielsen2007, 427]

### Examples

(@baure1) @danielsen2007, 428, example 144
```{r, echo=FALSE}
gloss_example(transliteration = "ro=kač-po-wo ro=pino-po",
             glosses = "3SG.M=go-PRFLX-COP 3SG.M=flee-PRFLX",
             annotation = "*rokačpow* *ropinop*",
             free_translation = "he (went away and) fled")
```

## Belep
### Details
Glottocode: nyal1254 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> In Belep, same-subject serialization with a limited set of verbs may be used to modify the aspect or modality of the clause. In such cases, the first verb provides modal information and the second verb provides lexical information. Switch-subject SVCs, by contrast, are used as a complementation strategy; in these instances, the first verb is the main verb (see §7.2.9). Belep SVCs are distinguished from compound verbs in that each verb in a SVC can serve independently as the main verb of a clause. [@mccracken2012, 370]

> The elements of the verb group always occur in a fixed order (they are an instance of templatic morphology; c.f. Bickel & Nichols 2007) and cannot undergo any syntactic operations except as a unit. Most elements in a verb group are not obligatory, but when they occur they have a conventionalized meaning and occur in an obligatory order. [@mccracken2012, 170]

> Intonation breaks, pauses, and repairs during the production of the verb group are most likely to occur immediately after the preface; however, these should be considered indicators of a phonological, rather than a grammatical, word boundary (Dixon and Aikhenvald 2002:24). [@mccracken2012, 330]

### Examples

Note that the glossing has been altered in \ref(@belep1) to reflect the grammatical dependency of *a=* 'NOM'.

(@belep1) @mccracken2012, 372, example 153
```{r, echo=FALSE}
gloss_example(transliteration = "te=mo may a=cama-la",
             glosses = "3SG.SBJ=live die NOM=father-3PL.POSS",
             free_translation = "Their father was about to die.",
             comment = "Te mo maac ya camala.")
```

(@belep1) @mccracken2012, 371, example 149
```{r, echo=FALSE}
gloss_example(transliteration = "ava=jara bae pê",
             glosses = "1PL.EXCL.SBJ=want bite bread",
             free_translation = "We want to eat the bread.",
             comment = "Ava jaar bae pê.")
```

## Bih
### Details
Glottocode: biha1246 <br> 
Family: Austronesian <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Blablanga
### Details
Glottocode: blab1237 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Bulo Stieng
### Details
Glottocode: bulo1242 <br> 
Family: Austroasiatic <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> Parmi ces critères [pour les CVS] - variables d'une langue à l'autre - les suivants ont été retenus pour être appliqués au stieng:
- le critère prosodique permettant de distinguer une CVS d'une concaténation de propositions... [@bon2014, 503]

Translation: Amongst the criteria [for SVCs] - which vary from language to language - the following have been retained for application to Stieng:
- the prosodic criterion which allows us to distinguish a SVC from a concatenation of clauses...

### Examples

> Dans l'example suivant (@bulo1), la séquence [sədiaŋ-dʔuar ɡɔk lot-kətaːŋ ʔuː rənaːp-rəsɨː] constitue un premier groupe de soufle, avec une prosodie montante. La séquence [ɓɑt joːh-ndiaŋ] constitue quant à elle un second groupe de souffle avec une intonation descendante. Ainsi, il est possible d'identifier les verbs *ɡɔk* 'ê.assis' *lot-kətaːŋ* 'ê.assis.à.genoux' comme faisant partie d'une série verbale au sein d'une même proposition; et *ɓɑt* 'plier' comme tête de la proposition suivant.
[@bon2014, 504-505]

Translation: In the following example (@bulo1), the sequence [sədiaŋ-dʔuar ɡɔk lot-kətaːŋ ʔuː rənaːp-rəsɨː] constitutes the first breath group with a rising contour. The sequence [ɓɑt joːh-ndiaŋ] in turn constitutes a second breath group with a falling intonation. Thus, it is possible to identify the verbs *ɡɔk* 'sit' *lot-kətaːŋ* 'kneel' as forming a serial verb construction within a single clause, and *ɓɑt* 'fold' as the head of the following clause.

(@bulo1) @bon2014, 505, example 839
```{r, echo=FALSE}
gloss_example(transliteration = "sədiaŋdʔuar ɡɔk lotkətaːŋ ʔuː rənaːprəsɨː (pause) ɓɑt joːhndiaŋ",
             glosses = "woman sit kneel on bamboo.mat (pause) fold shirt",
             free_translation = "The woman is kneeling on a bamboo mat and folding a shirt.")
```



## Burarra
### Details
Glottocode: bura1267 <br> 
Family: Maningrida <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

Note that glosses have been adapted for a general audience.

(@burarra2) @green1987, 79, example 5(lxxxii)
```{r, echo=FALSE}
gloss_example(transliteration = "Nginyi-pa gornabola ø-gana-ø ø-bamba-ø",
             glosses = "2MIN-DIR wallaby 2MIN.IMP-watch-IRR 2MIN.IMP-go_along-IRR",
             free_translation = "You go along watching for wallabies.")
```

(@burarra3) @green1987, 80, example 5(lxxxix)
```{r, echo=FALSE}
gloss_example(transliteration = "ngu-durtchi-nga ngu-bo-na",
             glosses = "1MIN-be_full-RL 1MIN-go-ANT",
             free_translation = "I'm full.")
```

(@burarra1) Green 1987:77, example 5(lxxviii)
```{r, echo=FALSE}
gloss_example(transliteration = "m-ba-nga a-rrik-nga gorrngunya",
             glosses = "3MIN>3MIN.mun-eat-PST 3MIN.an-crawl-PST grass",
             free_translation = "He crawled along eating grass.")
```

## Burmese
### Details
Glottocode: nucl1310 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasian

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> The sentence in (@burmese1) contains two verbs that are separated by the object of the second one, i.e. ‘hand’ [lɛʔ]. As suggested by the translation, we have here two clauses that are chained and not subordinated as in the previous example, while example (@burmese2) illustrated a SVC. Actually, in this sentence, the object of the second verb, i.e. ‘words’ [zəKa3], precedes the string of verbs, which suggests that the two verbs are closely associated syntactically. [@vittrant2012, 106]

<br>

> Vittrant (2006: 309): “Une construction de verbes en série (CVS) est une séquence de verbes qui ne sont syntaxiquement ni coordonnés ni subordonnés. Elle a d’autre part, la forme de surface d’un syntagme verbal unique.” [@vittrant2012, 106 footnote 9]

Translation: A serial verb construction is a sequence of verbes which are neither coordinated nor subordinated syntactically. It has, on the other hand, the surface form of a single verb phrase.

### Examples

(@burmese1) @vittrant2012, 107, example 5
```{r, echo=FALSE}
gloss_example(transliteration = "pəKan^3^ thɛ^3^ ye^2^ thɛ^1^ lɛʔ she^2^=Pi^3^",
             glosses = "plate inside water put hand wash=SUB.TPS",
             free_translation = "... after having poured water into the plate and having washed his hands...")
```

(@burmese2) @vittrant2012, 107, example 6
```{r, echo=FALSE}
gloss_example(transliteration = "θiN^2^bɔ^1^ pɔ^2^ =Ma^2^ θiN^2^bɔ^1^ pɔ^2^ =Ma^2^ zəKa^3^ la^2^ pyɔ^3^ =Tɛ^2^",
             glosses = "boat on LOC boat on LOC word come say REAL.ASS",
             free_translation = "On the boat, on the boat, (she) came to me to talk.",
             comment = "Lit. ...she came to tell me some words.")
```


## Buru
### Details
Glottocode: buru1303 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

