```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lingglosses)
#library(linguisticsdown)
#linguisticsdown::writeIPA()
```

## Ma'anyan!
### Details
Glottocode: maan1238 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@gudai1985, 284-286]

(@maanyan1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Mafea
### Details
Glottocode: mafe1237 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Mali
### Details
Glottocode: mali1284 <br> 
Family: Baining <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Manam
### Details
Glottocode: mana1295 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Mandarin Chinese
### Details
Glottocode: mand1415 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

> Indeed, the notion of coverb itself can be conceived of as a construction or recurring pattern rather than a part of speech or word class. In the coverbial construction, a verb forms a compound predicate together with the next verb, or MAIN VERB, which means that they share a subject, and that the coverbial phrase provides additional details about the event expressed by the main verb. [@wiedenhof2015, 134]

> Sentence (@mandarin2) will not be taken to mean (2) my writing of a letter on his behalf (‘I am writing a letter for him’) unless realized with a intonational break signaling an afterthought construction (§3.4): *Wó xiě xìn, gěi ta.* ‘I am writing a letter; for him, that is.’ But like (@mandarin1), this special prosodic pattern also allows for reading (1) ‘I am writing a letter; to him, that is.’. [@wiedenhof2015, 137]

### Examples

(@mandarin1) @wiedenhof2015, 136, example 5.34a
```{r, echo=FALSE}
gloss_example(transliteration = "wó gěi ta xiě xìn",
             glosses = "1.SG give 3 write letter",
             free_translation = "literally ‘I bestow on him and write a letter’ >",
             comment = "(i) ‘I am writing him a letter’ (the letter being written to him)
             (ii) ‘I am writing a letter for him’")
```

(@mandarin2) @wiedenhof2015, 136, example 5.34b
```{r, echo=FALSE}
gloss_example(transliteration = "wó xiě xìn gěi ta",
             glosses = "1.SG write letter give 3",
             free_translation = "literally ‘I write a letter and bestow on him’ >",
             comment = "‘I am writing a letter to him’")
```

(@mandarin3) @wiedenhof2015, 130, example 5.27
```{r, echo=FALSE, results='asis'}
gloss_example(transliteration = "Suóyi dà bùfen de rén <m> dōu shi yòng xīshì jié hūn <oh>",
             glosses = "therefore big part SUB person mm all be use western_style connect marriage oh",
             free_translation = "Therefore most people <mm> have a Western-style wedding <oh>.")
```




### Corpus Examples

## Mangghuer
### Details
Glottocode: minh1238 <br> 
Family: Mongolic-Khitan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

> In a main verb + auxiliary construction, the main verb is nonfinite. There are six morphological possibilities. In many instances, the main verb appears as just a verb stem, with no morphology at all. ...
An auxiliary verb always follows the main verb. The auxiliary verb (the final one, if more than one is present in a particular clause) bears finite morphology, or, if the clause is dependent, it is the auxiliary verb which bears whatever nonfinite morphology is used to define the relationship between the clause and another clause, on which it is dependent.[@slater2005, 132]

> In fact, it turns out that constructions of just these two types account for all of the appearances of nonfinite bare verb stems in Mangghuer. It would be possible to treat these two construction types as serializations, and this would certainly be commensurate with previous discussions of serial constructions. However, constructions of the first type can also be accounted for as instances of clause chaining, and we will see in this chapter that they fall naturally into the category of Mangghuer non-final clauses. Constructions of the second type have already been accounted for in Chapter 4, as verb + auxiliary constructions. Thus, although these Mangghuer constructions could be described as instances of serialization, I will describe them with other terms instead. [@slater2005, 231]

> No such postposing of zero-marked non-final clauses is possible. These clauses rely on their position (preceding another clause) to give them a semantic interpretation (preceding the action of the other clause in time). The fact that zero-marked clauses cannot be postposed thus appears to be not a syntactic fact, but rather, simply a semantic fact. If the zero marker had more semantic content, as do some of the other markers we have seen, zero-marked clauses might also be postposable. However, due to the lack of semantic specificity which obtains when no suffix at all is used, linear positioning must bear an unusual semantic load, and postposing becomes impossible. [@slater2005, 262]

### Examples

(@mangghuer1) @slater2005, 135, example 43
```{r, echo=FALSE}
gloss_example(transliteration = "nughuai pusa di ge-jiang.",
             glosses = "dog another eat do-OBJ:PERF",
             free_translation = "the dog also ate (it).")
```

(@mangghuer2) @slater2005, 244, example 50
```{r, echo=FALSE, results = 'asis'}
gloss_example(transliteration = "gan xi gan=ni bulai=ni liang-banjier cida danana gerdi kerla gher-gha-jiang.",
             glosses = "3SG go 3SG=GEN child=ACC two-piece tear after roof throw go_out-CAUS-OBJ:PERF",
             free_translation = "He (Monkey) went, tore his child into two pieces, (and) threw (the pieces) onto the roof.")
```

(@mangghuer3) @slater2005, 259, example 81
```{r, echo=FALSE}
gloss_example(transliteration = "zaohuang=du xi shuer=ni bari gui-jiang.",
             glosses = "kitchen=DAT go chopsticks=ACC take run-OBJ:PERF",
             free_translation = "(she) went into the kitchen, took the chopsticks and ran.")
```

## Mapudungun!
### Details
Glottocode: mapu1245 <br> 
Family: Araucanian <br>
Macroarea: South America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Citation [@smeets2008, 176; @zuniga2006, 10-11, 38]

(@language1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Mawng
### Details
Glottocode: maun1240 <br> 
Family: Iwaidjan <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

Although there is some evidence for the serial verb construction forming a grammatical unit [@singer2006, 132], there is no grammatical marking involved which could potentially signal the continuation of a syntactic unit across a pause.

### Examples

(@mawng1) @singer2006, 131, example 4-85
```{r, echo=FALSE, results='asis'}
gloss_example(transliteration = "Waka-pa k-i-wani-ø k-ani-mirlaju-n ja y-ara ja single.",
             glosses = "DIST.LAND-EMPH PRS-3MASC-sit-NPST PRS-3MASC/3LAND-wait_for-NPST MASC MASC-other MASC single_man",
             free_translation = "During that time he keeps on waiting (for a wife), the single man.")
```

(@mawng2) @singer2006, 131, example 4-86
```{r, echo=FALSE}
gloss_example(transliteration = "K-a-wani-ø apu-la-ngung walij",
             glosses = "PRS-3PL-sit-NPST 3PL/3EDIB-eat-PST.CONT food",
             free_translation = "They were eating food.")
```

(@mawng3) @singer2006, 132, example 4-87
```{r, echo=FALSE}
gloss_example(transliteration = "Nuka ja k-i-wani-ø k-ani-wurru-n.",
             glosses = "PROX.MASC MASC PRS-3MASC-sit-NPST PRS-3MASC/3LAND-know-NPST",
             free_translation = "He's the one who knows.")
```

## Maybrat-Karon
### Details
Glottocode: maib1239 <br> 
Family: Isolate <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Mehek
### Details
Glottocode: mehe1243 <br> 
Family: Sepik <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Menya!
### Details
Glottocode: meny1245 <br> 
Family: Angan <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Source [@whitehead2004, 29-30, 181-187]

(@menya1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Meriam
### Details
Glottocode: meri1244 <br> 
Family: Eastern Trans-Fly <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Meyah
### Details
Glottocode: meya1236 <br> 
Family: East Bird's Head <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Mian
### Details
Glottocode: mian1256 <br> 
Family: Nuclear Trans-New-Guinea <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Momu-Fas
### Details
Glottocode: fass1245 <br> 
Family: Baibai-Fas <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Mon
### Details
Glottocode: monn1252 <br> 
Family: Austroasiatic <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Mundari
### Details
Glottocode: mund1320 <br> 
Family: Austroasiatic <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Mungbam
### Details
Glottocode: abar1238 <br> 
Family: Bantoid <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Murrinhpatha
### Details
Glottocode: murr1258 <br> 
Family: Southern Daly <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: ?

### Examples

(@murrinhpatha2) @blythe2009, 364, On the Flat line 103
```{r, echo=FALSE}
gloss_example(transliteration = "ngunu-murrk-nu=ngi",
             glosses = "1SG.S.7go.FUT-eat-FUT=1SG.S.1sit.FUT",
             free_translation = "I want to eat some.")
```

(@murrinhpatha3) @blythe2009, 378, Ningbingi line 13
```{r, echo=FALSE}
gloss_example(transliteration = "me-nge-dha=pirri",
             glosses = "3SG.S.say.PSTIMP-3SG.IO-PST=3SG.S.3stand.PSTIMP",
             free_translation = "He was saying to her...")
```

(@murrinhpatha1) @blythe2009, 364, On the Flat line 87
```{r, echo=FALSE, results = 'asis'}
gloss_example(transliteration = "ku panda thanggugu thurdana-ngintha-yitj-tha=tjini ha ha",
             glosses = "animate that.you.know what_animate? 2SG.S.30.PSTIMP-DU.F.NSIB-tell_story-PST=2SG.S.1sit.PSTIMP ha ha",
             free_translation = "That thing of the 'ku' class, whatitsname. You two were telling that story. ")
```

## Mwotlap
### Details
Glottocode: motl1237 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

