
<!-- # ```{r setup, include=FALSE} -->
<!-- # knitr::opts_chunk$set(echo = FALSE, cache = TRUE) -->
<!-- # library(lingglosses) -->
<!-- # #library(linguisticsdown) -->
<!-- # #linguisticsdown::writeIPA() -->
<!-- # ``` -->

## Haida
### Details
Glottocode: nort2938 <br> 
Family: Isolate <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> It is necessary to carefully distinguish compound verbs from serial constructions (which do not occur in Haida) on the one hand, and from sequences of main verbs plus verb-complex elements (which are phrases rather than lexical items) on the other. Like all phrases and unlike verb-verb compounds, sequences of verb followed by verb-complex element(s) exhibit productivity limited only by semantic compatibility. ... In all languages with serial verbs, verb sequences describing acts of moving and then getting something have the verbs in that order (Durie 1997:304, 310, 335-336). Note, then, the order of verb and movement verb-complex element in \ref(@haida1) and \ref(@haida2), and also the order of the transient verb-complex elements vis a vis the head verb in \ref(@haida3). [@enrico2003, 1125-1126]

Note that I understand Enrico's argument for Haida not having SVCs to be that the order of verbal elements is unlike the SVCs discussed in @durie1997; however, this argument is orthogonal to the data that he presents.

(@haida1) @enrico2003, 1125, example 36b
```{r}
gloss_example(transliteration = "ya.aats' 'la tl'ak'a qaa-gya.ang-gang.",
             glosses = "knife 3 whet walk-PROG-PRS",
             free_translation = "He is walking along whetting a knife.")
```

(@haida2) @enrico2003, 1126, example 39a
```{r}
gloss_example(transliteration = "'laa-ga 'la di.ing dluwii-gwaang-gang.",
             glosses = "3-POSTP 3 search run-around-PRS",
             free_translation = "He is running around searching for her.")
```

(@haida3) @enrico2003, 1127, example 40b
```{r}
gloss_example(transliteration = "'la hlranggula 7in-gan.",
             glosses = "3 work leave_to_do-PST",
             free_translation = "He went to work (on a vehicle).")
```

## Hakhun Tangsa
### Details
Glottocode: noct1238 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> When we have multiple verbs in a row it is hard to tell whether we have multiple predications, and if so, where one ends and the other begins. ... There are a few formal means that help us delimit a predication. This involves morphological elements marking the boundary of the verb stems on both sides in two constructions – the auxiliary construction and the prohibitive construction. In addition, the position of certain verbs and knowledge of the possible semantic relations between the verbs can help us infer whether there are multiple predications and where the boundary is.  [@boro2017, 278]

> Thus, the checked stems occur in more nominalized constructions like the auxiliary construction, stand-alone nominalized sentences, or complement clauses. The open stems, on the other hand, function as finite verbs of finite clauses directly marked with the inflected operators (without an intervening auxiliary verb). [@boro2017, 276]

### Examples

(@hakhun1) @boro2017, 276, example 511
```{r}
gloss_example(transliteration = "vàn nɤ́ twè heʔ keʔ kuʔ t-iʔ",
             glosses = "fire LOC take keep go give PST-1PL",
             free_translation = "(We) throw (it) in the fire.")
```

(@hakhun2) @boro2017, 285, example 529
```{r}
gloss_example(transliteration = "pik ibə́ nâm ibə́ ŋé vù l-iʔ-mə",
             glosses = "paddy.field that plant.type that finish cut NF-1PL-NF",
             free_translation = "After finishing cutting the paddy field, the weeds...")
```

(@hakhun3) @boro2017, 289, example 533
```{r}
gloss_example(transliteration = "irə́kə́ bátənî mó rik nɤʔ t-ɤʔ",
             glosses = "there perhaps by.mistake die tread PST-1SG",
             free_translation = "There perhaps (I) trod (the ghost child) to death by mistake.")
```

<!-- ## Hano -->
<!-- ### Details -->
<!-- Glottocode: hano1246 <br>  -->
<!-- Family: Austronesian <br> -->
<!-- Macroarea: Papunesia -->

<!-- ### Coding -->
<!-- IU: 1 <br> Grouping: ? <br> Grammatical: 1 -->

<!-- > Citation [@source2001] -->

<!-- ### Examples -->

<!-- (@language1) @source2001, pg, example 27 -->
<!-- ```{r, echo=FALSE} -->
<!-- gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:", -->
<!--              glosses = "downward NAME-DAT shout say-PST", -->
<!--              free_translation = "I shouted it down to Solo:bu.") -->
<!-- ``` -->


## Haruai
### Details
Glottocode: haru1245 <br> 
Family: Piawi <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> While a Haruai clause can contain at most one morphologically marked verb, many Haruai clauses contain more than one verb, these are serial verb constructions... Note that in the Haruai serial verb construction, it is always the last verb that shows morphological marking. [@comrie1995, 28]

> There is one further observation that needs to be made concerning the interpretation of morphological marking on verbs. In the third person singular of the present tense, Haruai has a zero inflection. In positive declarative sentences, the verb is nonetheless marked, by the declarative suffix *-a*. In interrogative or negative sentences, however, there is no declarative suffix, and the verb thus appears overtly in the same form as nonfinal verbs in a serial verb construction, as in \ref(@haruai2). ... However, such third person present forms can always be distinguished from morphologically unmarked verb forms by changing the person-number... [@comrie1995, 29]

### Examples

(@haruai1) @comrie1995, 28, example 11
```{r}
gloss_example(transliteration = "An absö rag h-n-ng-a.",
             glosses = "we vegetables carry come-FUT-1PL-DECL",
             free_translation = "We will bring vegetables.")
```

(@haruai2) @comrie1995, 29, example 14
```{r}
gloss_example(transliteration = "Nöb'-akw ögapm p?",
             glosses = "man-that what take:PRES:3SG",
             free_translation = "What has that man taken?")
```

## Hatam
### Details
Glottocode: hata1243 <br> 
Family: Hatam-Mansim <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

> Hatam allows sequences of verbs without intervening conjunctions, usually called serial verb constructions. Although there are instances of three (or more) verbs, I will make my comments on sequences of two verbs. In most cases both verbs are marked for subject person/number... [@reesink1999, 97]

> When explicit conjunctions are employed it is clear that the sequence involves coordination of clauses. The same observation holds for sequences interrupted by a pause, as in example \ref(@hatam2). [@reesink1999, 98]

### Examples

(@hatam1) @reesink1999, 97, example 4.95
```{r}
gloss_example(transliteration = "Munggwom de di-ma i-pim i-nggum lo ig-yem hum ngum.",
             glosses = "child POS REL-that 3PL-cry 3PL-hungry but 3PL-eat fire coal",
             free_translation = "Those children are crying from hunger but they are eating charcoal.")
```

(@hatam2) @reesink1999, 97, example 4.97
```{r}
gloss_example(transliteration = "Lene i-ngot igy-o, i-com tut paku ni-kinei-o...",
             glosses = "then 3PL-tie house-or 3PL-throw along nail 3SG-bad-or",
             free_translation = "Then they built houses, they threw away nails that were bad...")
```

## Hidatsa
### Details
Glottocode: hida1246 <br> 
Family: Siouan  <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> There are two types of serial verb constructions in Hidatsa. In the first type (the majority of cases), except for the final verb in the sequence, all preceding component-verbs are marked with the contemporaneous suffix that triggers ablaut on the immediately preceding vowel. In the other type, the nonfinal verbs are not marked with the contemporaneous suffix and so do not ablaut. ... Subjects of verbs in serial verb constructions are coreferential and, with the exception of component-verbs that precede motion verbs in unmarked serial constructions, usually marked with pronominal prefixes. In the normal rate of speech serial verb constructions *usually* [emphasis NP] form a pitch phrase. [@park2012, 538-539]

> Contemporaneous serial verb constructions are always paraphrasable into two clauses that are then conceptualized as two separate events; i.e., replacing the contemporaneous suffix with the coordinative suffix yields two clauses that are conceptualized as two separate events (typically, events in sequence). [@park2012, 541]

> None of the elements in an unmarked serial construction are marked with ablaut-triggering contemporaneous suffixes if the semantics of the serialized construction encodes intent, desire, or purpose. The so-called purposive serial verb constructions anticipate future events, whereby the intended action follows the intent or motion encoded in the second verb. [@park2012, 544]

Note that the final verb in a clause *must* take an illocutionary suffix, which is non-zero except for imperatives to single subjects [@park2012, 224-225].

### Examples

(@hidatsa1) @park2012, 540, example 111
```{r}
gloss_example(transliteration = "miidí-Ø nuwí-Ø úˀsia-c",
             glosses = "crawl-CONT go.around-CONT arrive-DECL",
             free_translation = "He arrived by crawling.",
             annotation = "Miidá nuwá úˀsiaˀc.")
```


(@hidatsa2) @park2012, 545, example 124
```{r}
gloss_example(transliteration = "ma-iréˀ goowí-waa-c",
             glosses = "1POS-speak finish-1CAUS.DIR-DECL",
             free_translation = "I finished talking.",
             annotation = "Maréˀ goowíwaac.")
```

## Highland Populuca!
### Details
Glottocode: high1276 <br> 
Family: Mixe-Zoque <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> Serial verb constructions, especially nuclear serial verbs, make up a unique type of multi-verb construction in that, as stated above, more than one verb forms a complex, tightly bound unit requiring no special linking morphology that denotes a single event such that they share arguments, aspect, tense and mood. Each of these properties are described here.
 [@dejongboudreault2009, 674]

(@highland1) Aiton 2016:200, example 27
```{r}
gloss_example(transliteration = "a",
             glosses = "word",
             free_translation = "here")
```





## Hinuq
### Details
Glottocode: hinu1240 <br> 
Family: Nakh-Dagestanian <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> Serial verb constructions in Hinuq consist of two juxtaposed verbs which act together as a single predicate. The first verb appears only as a stem without any further suffixes but retains its agreement prefixes. The second verb takes all inflectional and derivational suffixes that apply to the whole complex. ... Many serial verbs belong to the semantic domain of verbs of motion and physical actions, but they cannot be characterized by other semantic criteria (e.g. they do not follow iconic principles, they do not denote culturally important concatenations of events). Serial verbs cannot be regarded as morphosyntactic devices (except for those expressing causative constructions). [@forker2013, 697]

> Nakh-Daghestanian languages do not belong to the languages that are generally known to employ verb serialization. Nevertheless, Hinuq serial verb constructions match well with Aikhenvald’s (2006) definition. They always consist of two verbs that together form a single predicate and have one single intonation contour. The two verbs are phonologically independent words, but they form one grammatical word, as can be seen by the argument structure and the agreement. Serial verbs share all arguments and modifiers and the agreement. [@forker2013, 702]

### Examples

(@hinuq1) @forker2013, 699, example 1283a
```{r}
gloss_example(transliteration = "hayɬoy ɣʷero-z t’oq hat͡ɬi-cat͡ɬi-š",
             glosses = "he.ERG cow-DAT knife bump-throw-PST",
             free_translation = "He repeatedly hit the cow with the knife.")
```
<!-- fix example -->

(@hinuq1) @forker2013, 700, example 1285b
```{r}
gloss_example(transliteration = "hibayɬi kek-Ø-ece goɬi polnostyu kek-Ø-ece goɬi de",
             glosses = "there break-I-melt be.IRR completely break-I-melt be.IRR 1SG",
             free_translation = "There I (masc.) would be smashed off, completely smashed off.")
```
<!-- fix example -->

## Ho
### Details
Glottocode: hooo1248 <br> 
Family: Austroasiatic <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

> We will see that Ho has nuclear serialization. The verbs in a nuclear serialization form a tight bond; there is only one set of arguments, and aspect marking modifies the complex event (rather than each verb root separately). ... We see that serialization is a productive feature of Ho grammar and, additionally, that many common V2s have grammaticalized to varying degrees. [@pucilowski2013, 202]

> In a serial verb construction, both verb roots form part of the same predicate i.e., there is no subordinating suffix, and they share subject clitics and object suffixes. Importantly, the negative particle and subject clitic – both of which must appear immediately before the main verb precede the entire serial verb construction, illustrating that it is a single unit. [@pucilowski2013, 203]

### Examples

(@ho1) @pucilowski2013, 204, example 6.55
```{r}
gloss_example(transliteration = "aɲ gaɽa-te iɖi-atu:-ɲ-pe",
             glosses = "1SG river-ALL take-leave-1SG-IMP.2PL",
             free_translation = "take me to the river and leave me")
```
<!-- fix examples -->

(@ho1) @pucilowski2013, 204, example 6.55
```{r}
gloss_example(transliteration = "naʔ=do miyaɖ baru_daru baʈi-gur-ke-ɖ-a, maʔ-gur-ke-ɖ-a=eʔ",
             glosses = "now=FOC one.INAN kusum_tree overturn-fall-PFV-TR-FIN cut_swing-fall-PFV-TR-FIN=3SG",
             free_translation = "now he brought the kusum tree down, [he] chopped it down")
```
<!-- fix examples -->

<!-- ## Huli -->
<!-- ### Details -->
<!-- Glottocode: huli1244 <br>  -->
<!-- Family: Nuclear Trans-New-Guinea <br> -->
<!-- Macroarea: Papunesia -->

<!-- ### Coding -->
<!-- IU: 1 <br> Grouping: ? <br> Grammatical: 1 -->

<!-- > Citation [@source2001] -->

<!-- ### Examples -->

<!-- (@language1) @source2001, pg, example 27 -->
<!-- ```{r, echo=FALSE} -->
<!-- gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:", -->
<!--              glosses = "downward NAME-DAT shout say-PST", -->
<!--              free_translation = "I shouted it down to Solo:bu.") -->
<!-- ``` -->


## Hup
### Details
Glottocode: hupd1244 <br> 
Family: Naduhup <br>
Macroarea: South America

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> The Hup verb compound is here understood as a type of serial verb construction that involves contiguous ‘nuclear’ serialization, resulting in a single phonological word ... In those Hup compounds that follow clear iconic principles, verb order corresponds to temporal succession of actions, including sub-events in a cause-effect relationship. Those Hup compounds that are less iconic (though perhaps not completely non-iconic, as discussed below) encode coincident motion, posture, and manner ... Finally, Hup verb compounds do not allow duplicate participant roles, such as two agents or two objects...
[@epps2008, 393]

> As a single-predicate, cosubordinate construction, a compound verb in Hup must encode an event that has a degree of conceptual unity. When this unity is not judged to be present, the events are necessarily represented by two predicates, linked by subordination or coordination strategies. However, among compound verbs, the linked stems can encode a variety of more or less conceptually integrated sub-events. [@epps2008, 395]

### Examples

(@hup1) @epps2008, 396, example 9
```{r}
gloss_example(transliteration = "ʔæ̌ytæn=yɨ́ʔ hɨd mæh-b’uy-d’əh-ham-yɨ́ʔ-ay-áh!",
             glosses = "together=TEL 3PL kill-throw-send-go-TEL-INCH-DECL",
             free_translation = "Together they killed them and threw them out!")
```


<!-- (@hup2) @epps2008, 397, example 15 -->
<!-- ```{r, echo=FALSE} -->
<!-- gloss_example(transliteration = "tɨ́h-ǎn kayǎk tɔ́ʔ hɨd g’ɔʔ-nɔ́ʔ-ɔ̃́y", -->
<!--              glosses = "3SG-OBJ manioc tuber 3PL pull_up-give-DYNM", -->
<!--              free_translation = "They’ve pulled up manioc and given it to her.") -->
<!-- ``` -->


Please note that not all diacritics could be represented accurately in (@hup3); please consult original source for original example.

(@hup3) @epps2008, 400, example 25
```{r}
gloss_example(transliteration = "tɨh yok-d’əh-ham-yæh-kamí=mah",
             glosses = "3SG poke-send-go-FRUST-time_of=REP",
             free_translation = "At the moment when she poked (the knife) in...")
```
<!-- fix examples -->

## Iatmul
### Details
Glottocode: iatm1242 <br> 
Family: Ndu <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

### Examples

(@iatmul1) @jendraschek2012, 419, example 19.39
```{r}
gloss_example(transliteration = "da'mage laavwi-di li'-ka-di'",
             glosses = "door open-3PL stay-PRS-3SG.MASC",
             free_translation = "They left the door open. (lit: 'they opened the door (and it) stays (open))")
```


## Imonda
### Details
Glottocode: imon1245 <br> 
Family: Border <br>
Macroarea: Papua New Guinea

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: ?

> Serialisation as described there consists of a lexical verb stem followed by one or more verb stems that are highly constrained and have mainly aspectual and other grammatical functions. The combination of these serialised verb stems is considered to constitute one word on the following grounds. First, there may be no pause between the stems and individual intonation contours of the stems is impossible. Second, the order of the serialised elements is strict. Third, the serialised stems share one set of verbal morphology and may not have individual arguments. [@seiler1985, 194]

> To conclude this section: Imonda possesses no morphosyntactic means to co-ordinate independent clauses. Clauses may be simply juxtaposed without any formal link. Frequently, juxtaposed clauses are reduced, often containing nothing but a verb. However, sequences of such reduced clauses are not to be regarded as monoclausal as they may have individual intonation contours, individual tense/mood marking, and individual core arguments. Semantically they do not form a unity (unlike the Vagala example cited above), but rather depict consecutive events. As far as the term verb serialisation is concerned, in this grammar this term is restricted to those cases where the individual stems form a tightly knit unity with one set of verbal morphology (5.41, and is not used for cases of reduced clauses that are neither semantically, syntactically nor phonologically tightly bound. Juxtaposed clauses need not necessarily involve a verb at all, in which case the term 'verb serialisation' would be obviously inappropriate. [@seiler1985, 196-197]

### Examples

(@imonda1) @seiler1985, 93, example 88
```{r, results='asis'}
gloss_example(transliteration = "tõgõ fe-lõh-fna; tëla-l-fa ha-pia-fna-ba, pe-m peha-fna",
             glosses = "thus do-stand-PROG husband-NOM-TOP DIR-come-PROG-TOP fear-CAUS go_down-PROG",
             free_translation = "They used to do it like this; when her husband came, she would run away in fear.")
```

(@imonda2) @seiler1985, 104, example 156
```{r}
gloss_example(transliteration = "kë-l tad-pada-hape",
             glosses = "bone-NOM CL-put-come_back",
             free_translation = "He put the bones there and came back.")
```

## Irabu
### Details
Glottocode: irab1238 <br> 
Family: Japonic <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> There are two major differences between compound verbs and agglutinative SVCs in Irabu. First, V2 in an SVC never undergoes sequential voicing, while V2 in a compound may do so ... Second, unlike compounds, V1 and V2 in an SVC constitute separate primary stems. Since each stem is a primary stem, each can carry its own derivational affix as long as it is semantically appropriate. As a single word, however, the serialised stems carry a single inflectional affix. [@shimoji2008, 302-303]

### Examples

(@irabu1) @shimoji2008, 303, example 6-71
```{r}
gloss_example(transliteration = "ibi-pazїmi-sїmi-tar",
             glosses = "plant-begin-CAUS-PST",
             free_translation = "ordered someone to begin planting")
```


## I'saka
### Details
Glottocode: kris1246 <br> 
Family: Sko <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> Independent verbs may be used in serial verb constructions, provided that they show identical subject agreement, and the same specifications for polarity, aspect, etc. Serial verb constructions are common constructions in the languages of New Guinea, and their use in I'saka can be seen in the following example, which shows both manner-of-motion and simple changeof-location verbs serialised together, each with their own inflection for subject, necessarily agreeing (if there is overt agreement – the 3PL subject in \ref(@isaka3) evokes no overt agreement on the verbs because these verbs take no agreement prefix for 3PL). [@donohuesanroque2004, 89]

> Adjacent verbs may sometimes share constituents, even when they do not form a serial verb construction. Object noun phrases are normally obligatory in bivalent clauses, but may be omitted if the clause immediately prior has the same object. Intonation patterns show that these are distinct clauses, rather than serial verbs. [@donohuesanroque2004, 88]

### Examples

(@isaka2) @donohuesanroque2004, 89, example 291
```{r}
gloss_example(transliteration = "A nuo k-aliè k-ele' ya.",
             glosses = "pig.MASC big 3SG.MASC-flee 3SG.MASC-go COMP",
             free_translation = "The big pig ran away.")
```

(@isaka3) @donohuesanroque2004, 89, example 292
```{r}
gloss_example(transliteration = "Dakau onde akane wíysau.",
             glosses = "children go hide forest",
             free_translation = "The children went and hid in the bush.")
```

(@isaka1) @donohuesanroque2004, 88, example 283
```{r}
gloss_example(transliteration = "Sù d-ai, n-òu.",
             glosses = "sago 1SG-get 1SG-eat",
             free_translation = "I got some sago, and ate (it).")
```

(@isaka4) @donohuesanroque2004, 92, line 6
```{r}
gloss_example(transliteration = "Pùng dina' tù, di makaing wii.",
             glosses = "y.sibling 1SG.POSS 3SG.NMASC.come 1SG.ACC help 3SG.NMASC.do",
             free_translation = "My little sister comes, she helps me.")
```

## Jamaican Creole English
### Details
Glottocode: jama1262 <br> 
Family: Indo-European <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> SVCs involve two or more verbs brought together without a complementizer, conjunction or infinitive marker, and with no pause. If TMA or negation are marked, the marking on all verbs agrees, and typically only occurs on the first. There is normally a single expressed subject, and one direct object (if any)... [@patrick2008]

> In Jamaican SVCs, it is also possible to focus the verbs in an SVC separately through predicate clefting ... [@hollington2015, 164]

### Examples

(@jcreole1) @hollington2015, 162, example 214
```{r}
gloss_example(transliteration = "Anansi iit kyari uom di pat",
             glosses = "Anansi eat carry home DEF pot",
             free_translation = "Anansi ate and carried the pot home.")
```

(@jcreole2) @hollington2015, 164, example 223
```{r}
gloss_example(transliteration = "A gi im bai di buk gi im",
             glosses = "COP give 3SG buy DEF book give 3SG",
             free_translation = "(S)he bought the book *for* him/her.")
```

(@jcreole3) @hollington2015, 164, example 223
```{r}
gloss_example(transliteration = "A bai im bai di buk gi im",
             glosses = "COP buy 3SG buy DEF book give 3SG",
             free_translation = "(S)he *bought* the book for him/her.")
```
