```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lingglosses)
#library(linguisticsdown)
#linguisticsdown::writeIPA()
```

## Kaba
### Details
Glottocode: kaba1281 <br> 
Family: Central Sudanic <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Kalam
### Details
Glottocode: kala1397 <br> 
Family: Nuclear Trans-New-Guinea <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Kam
### Details
Glottocode: kamm1249 <br> 
Family: Volta-Congo <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Kambera
### Details
Glottocode: kamb1299 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Karbi!
### Details
Glottocode: karb1241 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

### Examples

> Citation [@konnerth2017, 250-258]

(@karbi1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Karitiâna
### Details
Glottocode: kari1311 <br> 
Family: Tupian <br>
Macroarea: South America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?


> Occasionally in Karitiana two verbs may be serialized. In such cases, the second verb is often *tat*, "to go." The first verb is not similarly restricted. [@everett2006]

### Examples

(@karitiana1) @everett2006, 385, example 15.111
```{r, echo=FALSE}
#gloss_example(transliteration = "ɨ-kɨsepkɨsep ɨ-taka-tat-ø ɨ-taktang",
#             glosses = "1SG.ABS-jump 1SG.ABS-SAP-go-NFUT 1SG.ABS-swim",
#             free_translation = "I went jumping to swim.")
```



## Kayardild
### Details
Glottocode: kaya1319 <br> 
Family: Tangkic <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

> In (@kayardild2) the verbs belong to different clauses, as shown by the independent intonation contours (indicated here by a comma); a literal meaning is forced. ... In (@kayardild3) ... *thaatha* is a clausemate and follows the main verb, giving the 'go and return' meaning. [@evans1995, 308]

### Examples

(@kayardild1) @evans1995, 306-307, example 8-40
```{r, echo=FALSE}
gloss_example(transliteration = "nyingka dathin-ki dulk-i ngadi-ja kuluu-j, nyingka kalka-th",
             glosses = "2SG.NOM that-MLOC place-MLOC wound-ACT scratch-ACT 2SG.NOM sicken-ACT",
             free_translation = "You have despoiled that place by digging there, and you have fallen ill.")
```

(@kayardild2) @evans1995, 308, example 8-45
```{r, echo=FALSE}
gloss_example(transliteration = "niya kuujuu-j, thaa-th",
             glosses = "3SG.NOM swim-ACT return-ACT",
             free_translation = "He had a swim, and came back.")
```

(@kayardild3) @evans1995, 308, example 8-47
```{r, echo=FALSE}
gloss_example(transliteration = "niya kuujuu-ja thaa-th",
             glosses = "3SG.NOM swim-ACT return-ACT",
             free_translation = "He's gone for a swim (and will come back).")
```

## Keo
### Details
Glottocode: keoo1238 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> The following criteria are diagnostic of serial verb constructions in Kéo: <br>
1. A serial verb construction contains two or more verbs, which occur in the same clause describing a single event; <br> 
2. The whole serial verb complex is negated by a negator preceding the first verb (see 12.6.1);<br> 
3. All verbs come under the scope of one aspect or mood particle (see 12.2 and 12.3) which precedes the verb complex;<br> 
4. All verbs share a single subject, at the surface level;<br> 
5. The construction falls under the one intonation contour. That is, there are no pauses between the verbs in a serial construction. [@baird2002, 286-287]

> Based on auditory analysis, the verb complex falls under the one intonation contour. Therefore, in fluent speech there are no pauses with a grammatical function between contiguous verbs in serial verb constructions.<br>
Intonation does not fall nor rise on non-final verbs in a serial verb complex. A break in intonation, either in the form of a marked rise or fall on a non-final verb in a series, indicates that the verbs belong to different clauses. [@baird2002, 291]

### Examples

(@keo1) @baird2002, 292, example 11
```{r, echo=FALSE}
gloss_example(transliteration = "'Imu kai mbana pasa rédé So'a.",
             glosses = "3SG go walk market up So'a",
             free_translation = "She went to the market in So'a.")
```

(@keo2) @baird2002, 292, example 12
```{r, echo=FALSE}
gloss_example(transliteration = "'Imu kai, mbana pasa rédé So'a.",
             glosses = "3SG go walk market up So'a",
             free_translation = "She went, (she) went (walking) to the market in So'a.")
```

## Ket
### Details
Glottocode: kett1243 <br> 
Family: Teniseian <br>
Macroarea: Eurasian

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> While the use of asyndetic coordination to conjoin two coordinands at the level of noun phrases does not pose any problem, it is not the case with asyndetic coordination at the interclausal level. ... Ket verbs can stand on their own as independent sentences. Therefore, it is difficult to know whether two clauses are combined into a complex construction or rather constitute two standalone sentences linked in discourse. The criterion of intonation does not really seem to be of much help here. [@nefedov2015, 95]

### Examples

(@ket1) @nefedov2015, 107, example 4.51
```{r, echo=FALSE}
#gloss_example(transliteration = "dɨˀl du-k-hən du-den",
#             glosses = "child 3-TH-stand 3-weep",
#             free_translation = "The child is standing (and) crying.",
#             annotation = "dɨˀlʲ dúkkʌːn dúrɛn")
```


## Khatso
### Details
Glottocode: kats1235 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> Like many langauges in the Tibeto-Burman family, Khatso allows multiple verbs to occur adjacently in a single clause. These serial verbs work together to form a unitary predicate that expresses one, or sometimes two, events. Syntactically, these constructions, which may contain two or three verbs, share the same arguments and aspect markers, with no indication that one verb is subordinate to another. [@donlay2019, 273]

### Examples

(@khatso1) @donlay2019, 276, example 510
```{r, echo=FALSE}
gloss_example(transliteration = "ŋa31 ma24 vɤ323 kɯ33 li33.",
             glosses = "fish CL.GEN buy enter come",
             free_translation = "Buy (and bring) home a fish.")
```

(@khatso2) @donlay2019, 288, example 545
```{r, echo=FALSE}
gloss_example(transliteration = "tɛi33ni44 ta31 tɕo35 o31ma33tsa35 tɛi31 ko55 xa55 tsa31.",
             glosses = "this_way after then turnip_soup one CL.COL boil eat",
             free_translation = "Because of that, (we would) boil a batch of turnip soup to eat.")
```

## Khroskyabs!
### Details
Glottocode: guan1266 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

> Citation [@lai2017, 608-610]

(@khroskyabs1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```





## Konai
### Details
Glottocode: kona1242 <br> 
Family: East Strickland <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Korowai
### Details
Glottocode: koro1312 <br> 
Family: Nuclear Trans-New-Guinea <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Koyra Chiini Songhay
### Details
Glottocode: koyr1240 <br> 
Family: Songhay <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Krongo
### Details
Glottocode: kron1241 <br> 
Family: Kadugli-Krongo <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Kulsab
### Details
Glottocode: fait1240 <br> 
Family: Nuclear Trans-New-Guinea <br>
Macroarea: 

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Kumzari
### Details
Glottocode: kumz1235 <br> 
Family: Indo-European <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

> Kumzari has verb phrases with finite verbs, compound verbs, and auxiliary verbs. Compound verb phrases have morphosyntactic properties similar to those of finite verb phrases, but these are shared over two words: a deverb and a light verb. Auxiliary verbs do not necessarily share morphosyntax with their referent lexical verb, and they may be marked separately. [@walanonby2015, 112]

> The auxiliary verb takes the appropriate conjugation for the context, which is often the same as the lexical verb (@kumzari1). [@walanonby2015, 122]

### Examples

(@kumzari1) @walanonby2015, 122, example 271
```{r, echo=FALSE}
#gloss_example(transliteration = "brō ūn-ī pēna muẓgit.",
#             glosses = "go:2SG.IMP sit.IMP-2SG beside mosque",
#             free_translation = "Go sit beside the mosque.")
```

(@kumzari2) @walanonby2015, 122, example 273
```{r, echo=FALSE}
#gloss_example(transliteration = "č-im tkard–im ba xō kō-ō.",
#             glosses = "go:IMPF-1PL cultivate:IMPF-1PL for REFL mountain-the",
#             free_translation = "We will go cultivate for ourselves on the mountain.")
```

## Kunama
### Details
Glottocode: kuna1268 <br> 
Family: Isolate <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Kuot
### Details
Glottocode: kuot1243 <br> 
Family: Isolate <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Lachi
### Details
Glottocode: lach1248 <br> 
Family: Tai-Kadai <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Lamaholot
### Details
Glottocode: lama1277 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```



## Lamkang!
### Details
Glottocode: lamk1238 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

### Examples

> Citation [@thounaojamchelliah2007, 65-66]

(@lamkang1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```



## Lao
### Details
Glottocode: laoo1244 <br> 
Family: Tai-Kadai <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

> In referring to multi-verb constructions, I restrict this in general to sequences which normally form prosodically integrated units. [@enfield2007, 339]

> A multi-verb construction shows clause separability if it can be paraphrased with insertion of overt marking which forces a reading of the verbs as belonging to independent clauses, and where this causes no significant change in the basic semantic relationship between those verbs (although certain pragmatic effects may arise). 
One way to clause-separate a multi-verb construction is to insert between verbs a marked pause... [@enfield2007, 340]

### Examples

(@lao1) @enfield2007, 341, example 951a
```{r, echo=FALSE}
gloss_example(transliteration = "kap2-khùùn2 maa2 pathêêt4 hian2 tòò1",
             glosses = "go_back-return come country study continue",
             free_translation = "(They) came back to (their) country to continue (their) studies.")
```

(@lao2) @enfield2007, 482, example 1419
```{r, echo=FALSE}
gloss_example(transliteration = "ñaang1 paj3 qaw3 nam4 juu1 kakhuq1",
             glosses = "walk go take water be_at bucket",
             free_translation = "(He) walks (to) get water in a bucket.")
```

## Laven
### Details
Glottocode: lave1248 <br> 
Family: Austroasiatic <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Lavukaleve
### Details
Glottocode: lavu1241 <br> 
Family: Isolate <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Lega-Mwenga
### Details
Glottocode: lega1250 <br> 
Family: Bantu <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Lele
### Details
Glottocode: lele1270 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Lelepa
### Details
Glottocode: lele1267 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Lower Chehalis
### Details
Glottocode: lowe1427 <br> 
Family: Salishan <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

> In ɬəw̓á̓lməš frequently a series of predicates runs together as serial verbs, a phenomenon that has been discussed only a little in the Salish literature. ... In other words those sequences of verbs all are in the same clause, i.e. without conjunctions, complementizers, or prepositions joining them; they are identical in subject, aspect, tense and mood (and perhaps usually in voice). [@robertson2014, 147]

### Examples

(@lchehalis1) @robertson2014, 148, example 342
```{r, echo=FALSE}
#gloss_example(transliteration = "√ʔís-n √ʔúcət-w̓-n",
#             glosses = "√come-3.SBJ.IPFV √enter-INTR-3.SBJ.IPFV",
#             free_translation = "she come [and] enters")
```

(@lchehalis2) @robertson2014, 148, example 343
```{r, echo=FALSE}
#gloss_example(transliteration = "xʷát•√xʷətəq-aʔ √č̓ə́qʷ-ɬ-aʔ √pə́kʷ=čəp-aʔ-n",
#             glosses = "RDUP•√jump-IMP √get.up-INTS-IMP √ignite=fire-IMP",
#             free_translation = "hurry up, get up, and make a fire!")
```

