<!-- # ```{r setup, include=FALSE} -->
<!-- # knitr::opts_chunk$set(echo = FALSE, cache = TRUE) -->
<!-- # library(lingglosses) -->
<!-- # #library(linguisticsdown) -->
<!-- # #linguisticsdown::writeIPA() -->
<!-- # ``` -->

<!-- ## Cahuilla -->
<!-- ### Details -->
<!-- Glottocode: cahu1264 <br>  -->
<!-- Family: Uto-Aztecan <br> -->
<!-- Macroarea: North America -->

<!-- ### Coding -->
<!-- IU: 1 <br> Grouping: ? <br> Grammatical: 1 -->

<!-- > Citation [@source2001] -->

<!-- ### Examples -->

<!-- (@language1) @source2001, pg, example 27 -->
<!-- ```{r, echo=FALSE} -->
<!-- gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:", -->
<!--              glosses = "downward NAME-DAT shout say-PST", -->
<!--              free_translation = "I shouted it down to Solo:bu.") -->
<!-- ``` -->


<!-- ## Central Kanuri -->
<!-- ### Details -->
<!-- Glottocode: cent2050 <br>  -->
<!-- Family: Saharan <br> -->
<!-- Macroarea: Africa -->

<!-- ### Coding -->
<!-- IU: 1 <br> Grouping: ? <br> Grammatical: 1 -->

<!-- > Citation [@source2001] -->

<!-- ### Examples -->

<!-- (@language1) @source2001, pg, example 27 -->
<!-- ```{r, echo=FALSE} -->
<!-- gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:", -->
<!--              glosses = "downward NAME-DAT shout say-PST", -->
<!--              free_translation = "I shouted it down to Solo:bu.") -->
<!-- ``` -->


## Central Khmer
### Details
Glottocode: cent1989 <br> 
Family: Austroasiatic <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

### Examples

> Serial verbs are different: as far as one can tell, and with one single possible exception, all serial verbs are homophonous with currently attested main verbs. There are thus virtually no dedicated serial verbs. (Or if there are, we may question whether they are in fact verbs at all.) ... Serial verbs are in many cases indistinguishable from any asyndetically following clause. [@haiman2011, 270]

(@khmer1) @haiman2011, 276, example 62b
```{r}
gloss_example(transliteration = "ko: bamreah cenj pi: neum",
             glosses = "ox wriggle exit from yoke",
             free_translation = "The ox wriggled out from under the yoke.")
```

(@khmer2) @haiman2011, 280, example 69
```{r, results='asis'}
gloss_example(transliteration = "koat lu:k daj tev knong thawng.jiam haeuj daw:k luj muaj rial mau:k",
             glosses = "3 reach hand go in knapsack finish remove money one rial DIR",
             free_translation = "He reached into his knapsack and then removed a one-rial piece.")
```


## Chakali
### Details
Glottocode: chak1271 <br> 
Family: Volta-Congo <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

> Let us start by stating that the SVC in Chakali has the following properties: (i) a SVC is a sequence of verbs which act together as a single predicate, (ii) each verb in the series could occur as a predicate on its own, (iii) no connectives surface (coordination or subordination), (iv) tense, aspect, mood, and/or polarity are marked only once, (v) a verb involved in a SVC may be formally shortened, (vi) transitivity is common to the series, so arguments are shared (one argument obligatorily), (vii) the verbs in the series are not necessarily contiguous, and (viii) the grammar does not limit the number of verbs. [@brindle2017, 309]

### Examples

(@chakali1) @brindle2017, 310, example 20
```{r}
gloss_example(transliteration = "hÈmbÍÍ táwá fótò làgà dáá nÍ",
             glosses = "nail pierce picture hang wood POSTP",
             free_translation = "A picture hangs from a nail on a wooden pole.")
```


## Creek
### Details
Glottocode: cree1270 <br> 
Family: Muskogean <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

> Creek has several constructions in which an auxiliary verb is used with a main verb to express a stronger assertion, possibility, or aspect. ...the main verb is marked with durative -i: and agent person markers appear either on the auxiliary verb or on the main verb. [@martin2011, 298]

### Examples

(@creek1) @martin2011, 306, example 34
```{r}
gloss_example(transliteration = "nis-í: po:y-iphoy-í:-t ô:-s",
             glosses = "buy-DUR finish-SPON.IMPL-DUR-SS be.FGR-IND",
             free_translation = "They’ve bought it all.")
```

(@creek2) @martin2011, 306, example 38
```{r}
gloss_example(transliteration = "wana:y-í: mâ:h-ey-s",
             glosses = "tie.LGR-DUR keep.FGR-1SG.AG-IND",
             free_translation = "They’ve bought it all.")
```

## Daakaka
### Details
Glottocode: daka1243 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: 1 <br> Grammatical: 0

### Examples

> One fundamental criterion according to which all serial predicate constructions can be classified into one of two groups is finite marking: Depending on the non-initial predicate, some SPCs contain only one TAM marker (single marking), while in others, each predicate is preceded by its own TAM marker (multiple marking). Subject pronouns are never allowed in a non-initial predicate, the TAM marker always takes its monosyllabic form, as with third person singular subjects (see also section 5.1). [@vonprince2012, 304]

> The main two functions of intonation in Daakaka are to mark phrase boundaries and to distinguish between speech acts. [@vonprince2012, 37]

> Regarding multiple-marking SVCs, there is a tendency for every TMA marker to be aligned with a pitch rise: In three out of six cases, a high boundary tone was found at the position before the second tense marker (Figure 17). This is the predicted position if the two verbs in a multiple-marking SVC instantiate their own IPs. In two cases, a boundary tone was not found before the second TMA marker, and one case remains unclear. [@hopperdietzel2020, 138]

(@daakaka1) @vonprince2012, 308, example 6c
```{r}
gloss_example(transliteration = "mwe tavya te polo usili ló-ó swa vyan milye",
             glosses = "RL get_up CONJ climb follow plant-coconut one go on_top",
             free_translation = "she got up and climbed up a coconut palm")
```

(@daakaka2) @vonprince2012, 336, example 80a
```{r}
gloss_example(transliteration = "Ko=m en seaa ada mubuo mo nok?",
             glosses = "2SG=RL eat all CL2:1DU.INCL.POSS meat RL finish",
             free_translation = "Have you eaten all our meat?")
```

## Dagik
### Details
Glottocode: dagi1241 <br> 
Family: Narrow Talodi <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

> This section investigates verbal sequences (VS) in independent clauses. They have the following characteristics: (i) they consist of at least two verbs, (ii) aspect-mood is only marked once for the whole verbal sequence, (iii) the subject is only expressed once for the whole verbal sequence, and (iv) the verbal sequence can only be negated as a whole. These characteristics show that verbal sequences are monoclausal. The first verb of the sequence (V1) carries aspect-mood markers and agrees with the subject. The subsequent verb(s) is/are in status 2 (see section 4.7). [@vanderelst2016, 213]

### Examples

(@dagik1) @vanderelst2016, 214, example 69
```{r}
gloss_example(transliteration = "a-ŋɪ b-a=nda ðɛŋg-ɔ baɽ-a",
             glosses = "REF-1SG CL-COP=afterwards stay-FV go_out-FV",
             free_translation = "I will go out after a little while. (lit.: I will stay [and] go out)")
```

(@dagik2) @vanderelst2016, 216, example 83
```{r}
gloss_example(transliteration = "ŋ-əmasɛ ŋ-a ɾas-ɔ nag-a ð-ʊɽa",
             glosses = "CL-child CL-COP come-FV keep-FV CL-stealing_activity",
             free_translation = "'The child will become a thief.' (e.g. as a grown-up) (lit.: 'the child will come keep stealing')")
```

## Darma
### Details
Glottocode: darm1243 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

> When a verb functions as an auxiliary, it is found following another verb stem; the auxiliary bears the inflectional morphology of the utterance. In these cases, the two verbs form a single unit. In the following subsections I will outline the verbs found in combination. (These constructions may be better analyzed as serial verbs.) [@willis2007, 341]

> A verb includes a verb stem plus inflectional morphology. The stem is the morpheme that contains the lexical information of the verb. Verbs can be inflected for person, number, tense, aspect and mood. Verb stems generally do not appear alone except when they precede an inflected verb to form a complex verb. [@willis2007, 327]

### Examples

(@darma1) @willis2007, 342, example 36
```{r}
gloss_example(transliteration = "ning gu dar’umpha wan r’a-hen",
             glosses = "1PL POSS door.at reach come-1PL.NPST",
             free_translation = "(We) reach our door.",
             annotation = "ning gu dar’umpha wan r’ahen..")
```

(@darma2) @willis2007, 344, example 43
```{r}
gloss_example(transliteration = "hã yak to dee-nu ni-n-su",
             glosses = "then yak buy go-NOM AUX.EQ-1PL-PST",
             free_translation = "Then, (we) used to go buy a yak.",
             annotation = "hã yak to deen ninsu.")
```

## Dazaga
### Details
Glottocode: daza1242 <br> 
Family: Saharan <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> ...the two verbs in an SVC (as identified by various syntactic tests) are never separated by a coordinator, whereas clausal coordination is never asyndetic; the two verbs in an SVC always share at least one argument (as is typical in SVCs; cf. Kroeger 2004:229; Haspelmath 2016:309), whereas verbs in coordinate clauses need not share any arguments. [@walters2016, 216]

> The test of whether only one half or part of a construction can be questioned does not distinguish SVCs from coordinate constructions in Dazaga. A single constituent can be questioned from ... SVCs, as in \ref(@dazaga1) and \ref(@dazaga2). [@walters2016, 221]

### Examples

(@dazaga3) @walters2016, 222, example 585
```{r}
gloss_example(transliteration = "búrú X-lôn-X ànìSí X-dÍR-X",
             glosses = "hole 3.OBJ-dig.IMV-2 pure_sand 3.OBJ-take_out.IMV-2",
             free_translation = "Dig a hole to take out pure sand.",
             annotation = "búrú lôn ànìSí dÍR")
```

(@dazaga1) @walters2016, 221, example 579
```{r}
gloss_example(transliteration = "ínní ló-X-j ànìSí X-j-tíR",
             glosses = "what dig-3.OBJ-3 pure_sand 3.OBJ-3-take_out",
             free_translation = "What did he dig to take out pure sand?",
             annotation = "ínní lóì ànìSí díRÙ")
```

(@dazaga2) @walters2016, 221, example 580
```{r}
gloss_example(transliteration = "búrú ló-X-j ínní X-j-tíR",
             glosses = "hole dig-3.OBJ-3 what 3.OBJ-3-take_out",
             free_translation = "What did he dig a hole to take out?",
             annotation = "búrú lóì ínní díRÙ")
```

## Desano
### Details
Glottocode: desa1247 <br> 
Family: Tucanoan <br>
Macroarea: South America

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> As in other Eastern Tukanoan languages, SVCs are a highly productive morphological process in Desano. The SVCs in Desano can be characterized as being ‘continuous’ and ‘incorporating’; showing both symmetrical and asymmetrical constructions, and having both single and concordant marking. Concordance marking is not very common. [@delimasilva2012, 235]

### Examples

(@desano1) @delimasilva2012, pg235-236, example 58
```{r, results='asis'}
gloss_example(transliteration = "yuu-~dʉ yai—~yeko—~beda ~bibi wʉ-bida—~da i—~yo—~da",
             glosses = "one-CLS:day heron-grandmother-COM hummingbird fly-play-3PL.ANIM.PERF do-HSAY-3PL:ANIM.PERF",
             free_translation = "One day, the tuiuiu bird and the hummingbird were flying and playing.",
             annotation = "yuhunʉ̃  yahiyẽhkõmẽɾã mĩmĩ wʉbiɾaɾã iyõɾã")
```
<!-- fix example -->

(@desano2) @delimasilva2012, pg236, example 60
```{r}
gloss_example(transliteration = "pe-bʉ kudi-bʉ yʉʉ—~sa",
             glosses = "listen-NON3.PERF walk-NON3.PERF 1SG-ADD",
             free_translation = "I also listened and walked.",
             annotation = "peebʉ kuɾibʉ yʉ'ʉsã.")
```
<!-- fix example -->


## Dhao
### Details
Glottocode: dhao1237 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

> Juxtaposition refers to coordination without an overt linker. This type of construction occurs either on word, phrase, or clause level. Since there is no overt marking, intonation is the only means to identify conjoined units (Haspelmath, 2007: 7). [@balukh2020, 230]

> Paratactic constructions also appear like SVCs in terms of the consecutive occurrence of verbs. An extreme sequence of verbs is shown in \ref(@dhao2) below. It is a complement construction marked by the complementizer *na* (see §6.3.2.1). The matrix clause itself contains a SVC consisting of two verbs, *pèci* ‘to throw’ and the inflected verb *mere* ‘to take’. Furthermore, the complement clause has seven verbs that occur consecutively ... It is difficult to determine SVCs in this construction, as there are no overt syntactic markings. Contributing to this difficulty is the fact that subject and object deletion is a common pattern in natural discourse in Dhao. As such, argument sharing is difficult to identify. After looking at the construction in detail, it appears that it has two separate clauses tied together. The first clause involves V1-V3, while the second clause involves V4-V7. It is shown that they have separate subjects, even though the different subjects refer to the same referent. ... the complement clause covers only the first clause in this construction, while the second clause is a separate clause that designates another follow-up event, which is the event that takes place after another event (fetch). ... There still are two sequence of verbs, *lami madhutu* ‘go to fetch’ and *la’e tenge* ‘go to look for’. In turn, these will not be considered as SVCs either, due to their predicate-argument relation. As the result, no sequence of the seven verbs qualifies as a SVC. [@balukh2020, 259]

### Examples

(@dhao1) @balukh2020, 231, example 14
```{r, results='asis'}
gloss_example(transliteration = "k-ore doi, la-ku hèli èi na'i mea mèdi, mai nasu hèngu ne'e, te ja'a doi aad'o nga",
             glosses = "1SG-take money go-1SG buy water tobacco red black come boil yarn PROX.SG because 1SG money be_absent PART",
             free_translation = "I can earn money and I go to buy red and black dye and then I come to boil these yarns, because I don't have money.")
```

(@dhao2) @balukh2020, 259, example 116
```{r, results='asis'}
gloss_example(transliteration = "pèci m-ere na mai la-mi madhutu mai hia la-'e tenge",
             glosses = "throw 2PL-take COMP come go-2PL follow come give go-3SG look_for",
             free_translation = "after throwing it, (you) go home and took (him), (you) come to ask (him) to find...")
```

## Dla
### Details
Glottocode: dera1245 <br> 
Family: Senagi <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 1

> Non-finite verb stems are used in non-finite chain verbs (§7.3.1) and verbal nouns (§7.3.2). Serial verb constructions in non-finite chain clauses have a string of non-finite verb stems (with no cross-reference suffix in between) plus other affixes. ... Serial verb constructions in independent, subordinate and chain clauses can consist of strings of finite verb root plus cross-reference suffix(es) and other affixes. [@desousa2006, 261]

### Examples

(@dla1) @desousa2006, 447, example 7-70
```{r}
gloss_example(transliteration = "ser-i fa-hya-a numb-a,",
             glosses = "eat-1SG COMPL-1SG-3SG.FEM.OBJ SEQ-1SG",
             free_translation = "I ate, and after that I...")
```

(@dla2) @desousa2006, 38, example 1-4
```{r}
gloss_example(transliteration = "hwama-i fa-i-hi,",
             glosses = "hang-3SG.MASC leave-3SG.MASC-SIM",
             free_translation = "While he hang and left him there...")
```
<!-- ## Dongwang Tibetan -->
<!-- ### Details -->
<!-- Glottocode: kham1282 <br>  -->
<!-- Family: Trans-Himalayan <br> -->
<!-- Macroarea: Eurasia -->

<!-- ### Coding -->
<!-- IU: 1 <br> Grouping: ? <br> Grammatical: 1 -->

<!-- > Citation [@source2001] -->

<!-- ### Examples -->

<!-- (@language1) @source2001, pg, example 27 -->
<!-- ```{r, echo=FALSE} -->
<!-- gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:", -->
<!--              glosses = "downward NAME-DAT shout say-PST", -->
<!--              free_translation = "I shouted it down to Solo:bu.") -->
<!-- ``` -->

## Doromu-Koki
### Details
Glottocode: doro1266 <br> 
Family: Manubaran <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> While serial verb constructions do occur in Doromu-Koki, they are not nearly as productive as in other Papuan languages. In fact, they are very rare. They are limited to the verbs *afe* ‘return’, *baku* ‘find’, *bo* ‘go up’, *fere* ‘leave’, *ne* ‘go down’, *ni* ‘say’, *moi* ‘get’, *re* ‘do’, and *usi* ‘follow’, which indicates that they are usually tied to motion, combining two separate actions together into one event. [@bradshaw2012, 101-102]

### Examples

An single intonation contour is given in the original source for \ref(@doromukoki1).

(@doromukoki1) @bradshaw2012, 36, example 9
```{r, results='asis'}
gloss_example(transliteration = "Rei mona maka de re-yo ma yaku bura ri-yo urusa rofu.",
             glosses = "do properly only NEG do-2/3SG.PST and DM expose make-2/3SG.PST night BEN",
             free_translation = "He did not do it properly and so he was exposed at night.")
```

(@doromukoki2) @bradshaw2012, 102, example 113
```{r}
gloss_example(transliteration = "Maruo ve-sa dada ya ferei bai-yaka.",
             glosses = "movie see-2SG.PRS so 2 leave come-1SG.PST",
             free_translation = "You are watching a movie, so I left you and came.")
```

## Duhumbi
### Details
Glottocode: chug1252 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> A Duhumbi serial verb construction is defined as a minimum of two verb roots forming a predicate in any non-finite or finite context that refer to what can be construed as a single event (though with sub-events possible) and form a single phonosyntactic unit. [@bodt2020, 437]

> The modifying verb roots in asymmetrical SVCs can be compared to what have been called ‘converbs’ in other languages (e.g. Coupe 2007 for Mongsen Ao; Hyslop 2017 for Kurtöp). Their omission from the string or breaking up the string by adding non-finite markers to the individual roots, changes the entire meaning of the clause. [@bodt2020, 444]

### Examples

(@duhumbi1) @bodt2020, 438, example 655
```{r}
gloss_example(transliteration = "hentʰoŋ-kʰo pʰargiŋ kej oloʔ lek woj-ʨʰa-kʰen ose ʥu-loŋ=beʔ=oloʔ",
             glosses = "neck-LOC yoke place then field plough-eat-AGN like_that stay-PRF=COP.EX=then",
             free_translation = "(He) placed the yoke on the neck. And like that (the bull) had stayed (as) the one ploughing and eating the field.")
```
<!-- fix example -->

(@duhumbi2) @bodt2020, 446, example 668
```{r}
gloss_example(transliteration = "ɕapa kʰonow ɲis=gi lon-le-da, ‘Kʰar-han-ɲi’ in-ba=le.",
             glosses = "butcher siblings two=TOP come-do-IPFV call-bring-IMP say-NOM=COP",
             free_translation = "After those two butcher brothers came, (the demoness) said 'Call them over!'.")
```
<!-- fix example -->


## Duna
### Details
Glottocode: duna1248 <br> 
Family: Isolate <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: ?

> For Duna, I use the term ‘verb series’ (and related terms) to refer to sequences of verb roots for which the final verb in the sequence (Vlast) is inflected, bearing medial or final verb morphology, and the other verbs in the sequence (Vn) are uninflected. [@sanroque2008, 227]

### Examples

(@duna1) @sanroque2008, 113, example 84a
```{r}
gloss_example(transliteration = "no waya sa-ya ngo.",
             glosses = "1SG wash take-DEP go.PFV",
             free_translation = "I went to wash.")
```


## Dzongkha
### Details
Glottocode: dzon1239 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> The difference between the phonological word and the grammatical word can be illustrated in the clause *d'orzü=ci tâng ta-ge* 'Let me tell you a shortened version' (S16-1). The sequence *tâng ta-ge* is an example of a serial verb construction (SVC), and it consists of two grammatical words, *tâng* 'send' and *ta-ge* 'look-ADH'. Figure 46 shows, however, that the two grammatical words are part of a single phonological word. [@watters2018, 138]

### Examples

(@dzongkha1) @watters2018, 261, example 111
```{r}
gloss_example(transliteration = "zai numo-i gari jo dö ong mâ-lê phâ",
             glosses = "EX night-EMPH vehicle go sit PROB down=ABL over_there",
             free_translation = "Yikes! Vehicles even travel at night, down across there (to Kalimpong).")
```
