<!-- # ```{r setup, include=FALSE} -->
<!-- # knitr::opts_chunk$set(echo = FALSE, cache = TRUE) -->
<!-- # library(lingglosses) -->
<!-- # #library(linguisticsdown) -->
<!-- # #linguisticsdown::writeIPA() -->
<!-- # ``` -->


## N|uuki
### Details
Glottocode: nuuu1241 <br> 
Family: Tuu <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

> A serial verb construction in N|uuki is where one verb immediately follows another. Such a construction is aso referred to as a verbal compound. ... [N]othing can intervene between the two members of a serial verb construction. [@collinsnamaseb2011, 53]

### Examples

(@nuu1) @collinsnamaseb2011, 53, example 4a
```{r}
gloss_example(transliteration = "n-a si lʔama doŋki-si",
             glosses = "1SG-DECL IRR buy donkey-SG",
             free_translation = "I will buy the donkey.")
```

(@nuu2) @collinsnamaseb2011, 53, example 8c
```{r}
gloss_example(transliteration = "n-a kx'u-|x'õa-a ku",
             glosses = "1SG-DECL make-hunt-PFV 3SG",
             free_translation = "I made him hunt.")
```


<!-- ## Namuyi -->
<!-- ### Details -->
<!-- Glottocode: namu1246 <br>  -->
<!-- Family: Trans-Himalayan <br> -->
<!-- Macroarea: Eurasia -->

<!-- ### Coding -->
<!-- IU: 1 <br> Grouping: ? <br> Grammatical: 1 -->

<!-- > Citation [@source2001] -->

<!-- ### Examples -->

<!-- (@language1) @source2001, pg, example 27 -->
<!-- ```{r, echo=FALSE} -->
<!-- gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:", -->
<!--              glosses = "downward NAME-DAT shout say-PST", -->
<!--              free_translation = "I shouted it down to Solo:bu.") -->
<!-- ``` -->

## Ncane-Mungong
### Details
Glottocode: ncan1245 <br> 
Family: Bantoid <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Sentences with multiple verbs are common in Nchane. Those which consist of clauses of equal status could possibly be described as serial verb constructions (henceforth SVCs)... There are no morphological or syntactic markers of clausal subordination or of support clause status, nor of clausal adjunction in either example. Subject marking is usually associated with each verb of a sequence. ... the tense-marking strategy of the language [] usually establishes tense at discourse boundaries, with little or no tense marking for later sentences. [@boutwell2020, 284-285]

> Prosodies represent another possible indicator to differentiate SVCs. However, prosodic elements in Nchane sentences are somewhat inconsistently realized. Clause boundaries are often marked by pauses, but this is not always the case. Example \ref(@ncane1) presents multiple verbal elements with no apparent pause, change in intensity or tonal phenomena to mark a clause boundary. However, the two verbs may not be interpreted as encoding a single complex event, but rather two separate events in sequence. Thus, I interpret this sentence as comprised of two juxtaposed, independent clauses. ... Conversely, some sentences with multiple verbs which seem to express a single event, and thus, representative of an SVC analysis, have clear pauses following the verbs. Such is the case in \ref(@ncane2), which has a distinct pause following the verb ‘go’. [@boutwell2020, 286]

### Examples

(@ncane1) @boutwell2020, 286, example 13.3
```{r}
gloss_example(transliteration = "wɔ́ gɛ̄: à màŋ-kàlà mā-nē lē, wɔ́ ɲá wù lē",
             glosses = "2SG.FUT put in c6a-cassava.puff c6a-PROX APPL 2SG.FUT give 3SG APPL",
             free_translation = "...you put [it] (the poison) in this cassava puff and give [it] to him.")
```

(@ncane2) @boutwell2020, 287, example 13.4
```{r}
gloss_example(transliteration = "mɛ́ ń-tú ŋ́-gɛ̰᷆́: / ɲ̄-jí bvū-lɛ̰̀́: bɛ́ ɲàŋ yī shi",
             glosses = "1SG.PRO 1SG-then 1SG-go {} 1SG-eat c14-food with c9.meat c9AM c9.chicken",
             free_translation = "I then went and ate fufu with chicken...")
```


## Ndebele
### Details
Glottocode: nort2795 <br> 
Family: Bantu <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

Question words must occur in second position. Verbs in serial verb constructions do not occur as a single unit in first position; rather, one verb precedes the question word and one follows (field notes).

### Examples

(@ndebele1) @bowernlotridge2002, 82
```{r, results='asis'}
gloss_example(transliteration = "u-ngi-khangela-e u-bona-e e-ngi-ku-enza-yo lawa futhi u-ku-enza",
             glosses = "2SG-1SG-look_at-SUBJ 2SG-watch-SUBJ REL-1SG-CL15-do-REL that.CL3 also 2SG-CL15-do",
             free_translation = "Watch me,what what I do and you do it too.",
             annotation = "Ungkhangele ubone enkwenzayo lawa futhi ukwenza.")
```

<!-- ## Nese -->
<!-- ### Details -->
<!-- Glottocode: nese1235 <br>  -->
<!-- Family: Austronesian <br> -->
<!-- Macroarea: Papunesia -->

<!-- ### Coding -->
<!-- IU: 1 <br> Grouping: ? <br> Grammatical: 1 -->

<!-- > Citation [@source2001] -->

<!-- ### Examples -->

<!-- (@language1) @source2001, pg, example 27 -->
<!-- ```{r, echo=FALSE} -->
<!-- gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:", -->
<!--              glosses = "downward NAME-DAT shout say-PST", -->
<!--              free_translation = "I shouted it down to Solo:bu.") -->
<!-- ``` -->

## Ngiti
### Details
Glottocode: ngit1239 <br> 
Family: Central Sudanic <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> The Perfective Present is also used in serial verb constructions. The first verb in a serial verb construction is inflected for tense/aspect. Any following verb is found in the Perfective Present and carries a subject clitic which is often (but not necessarily) identical to the subject of the initial verb. Both negative markers ... are attested with the Perfective Present when used as second or subsequent verb in a serial verb construction. [@kutschlojenga1994, 253]

> This means that, when used as a consecutive, the Perfective Present does not refer to a particular time by itself, but that it indicates a continuation of the previously-set time of the narration. [@kutschlojenga1994, 254]

Note that there is no suffix associated with the Perfective Present [@kutschlojenga1994, 251].

### Examples

Please note that not all diacritics could be rendered faithfully and that the original source should be consulted.

(@ngiti1) @kutschlojenga1994, 224
```{r}
gloss_example(transliteration = "ma m-òzè m-ozò kapʉ̀",
             glosses = "1SG 1SG-want:PFV.PRS 1SG-plant:PFV.PRS bean",
             free_translation = "I want to plant beans.")
```

(@ngiti1) @kutschlojenga1994, 380
```{r}
gloss_example(transliteration = "k-òzè ná nd-ɔ̀nyʉ̀ tsìtsì",
             glosses = "3SG-want.PFV.PRS Q 3SG.LOG-eat:PFV.PRS banana",
             free_translation = "Does he want to eat a banana?",
             annotation = "kòzè ná ndÒnyÙ tsìtsì?")
```

<!-- ## Northern Qiang -->
<!-- ### Details -->
<!-- Glottocode: nort2722 <br>  -->
<!-- Family: Trans-Himalayan <br> -->
<!-- Macroarea: Eurasia -->

<!-- ### Coding -->
<!-- IU: 1 <br> Grouping: ? <br> Grammatical: 1 -->

<!-- > Citation [@source2001] -->

<!-- ### Examples -->

<!-- (@language1) @source2001, pg, example 27 -->
<!-- ```{r, echo=FALSE} -->
<!-- gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:", -->
<!--              glosses = "downward NAME-DAT shout say-PST", -->
<!--              free_translation = "I shouted it down to Solo:bu.") -->
<!-- ``` -->


## Northern Yukian
### Details
Glottocode: yuki1243 <br> 
Family: Yuki-Wappo <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Non-final verbs can be bare verb roots or can be verb roots affixed with derivational morphology, in Yuki serial verb constructions. The non-final verbs in the serial verb construction are never marked for tense or with *=mil*. The final verb in the sequence will be affixed with finite *=mil*. [@balodis2011, 665]

### Examples

(@yuki1) @balodis2011, 666, example 83
```{r}
gloss_example(transliteration = "sąmi ki-mas-i hąša hąp šuˀ kop-wok-tl=mil tat-k-il",
             glosses = "and:then DST-DSTR-ANIM again sing sit feather-dance-TR=FIN fix-PNCT-MPSV",
             free_translation = "And then in turn these others sing, sit, and dance the feather dance and fix themselves up.",
             annotation = "sámi: kimáse hąšáˀ hąp šú: kopwóktlmil tá:tkí:li.")
```

(@yuki2) @balodis2011, 667, example 84
```{r}
gloss_example(transliteration = "siˀi lak-t nąw-k-il=mil ki ˀipsak",
             glosses = "then:HSAY go.out-INTR see-PNCT-MPSV=FIN DST boy",
             free_translation = "Then going out, the boy looked.",
             annotation = "seˀéy lákti nąwkílmil kí ˀipsák")
```

<!-- ## Nungon -->
<!-- ### Details -->
<!-- Glottocode: yaum1237 <br>  -->
<!-- Family: Nuclear Trans-New-Guinea <br> -->
<!-- Macroarea: Papunesia -->

<!-- ### Coding -->
<!-- IU: 1 <br> Grouping: ? <br> Grammatical: 1 -->

<!-- > Citation [@source2001] -->

<!-- ### Examples -->

<!-- (@language1) @source2001, pg, example 27 -->
<!-- ```{r, echo=FALSE} -->
<!-- gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:", -->
<!--              glosses = "downward NAME-DAT shout say-PST", -->
<!--              free_translation = "I shouted it down to Solo:bu.") -->
<!-- ``` -->


## Nupe
### Details
Glottocode: nupe1254 <br> 
Family: Volta-Congo <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

### Examples

> A second syntactic argument concerns the fact that in all Nupe SVCs, only the initial verbal occurrence may be repeated ... [@kandybowicz2008, 54]

> Subject and object extraction from SVCs, for example, is permissible, as shown in \ref(@nupe3). [@kandybowicz2008, 56]

(@nupe1) @kandybowicz2008, 54, example 6a
```{r}
gloss_example(transliteration = "Musa du eci du kún.",
             glosses = "Musa cook yam cook sell",
             free_translation = "Musa DID IN FACT cook a yam and (then) sell it.")
```

(@nupe2) @kandybowicz2008, 54, example 6b
```{r}
gloss_example(transliteration = "Musa è fo èwò fo li.",
             glosses = "Musa PRS wash garment wash be_clean",
             free_translation = "Musa IS IN FACT washing the garment clean.")
```

(@nupe3) @kandybowicz2008, 56, example 9a
```{r}
gloss_example(transliteration = "Musa du eci kún o.",
             glosses = "Musa cook yam sell FOC",
             free_translation = "MUSA cooked a yam and (then) sold it.")
```

(@nupe4) @kandybowicz2008, 54, example 9d
```{r}
gloss_example(transliteration = "Èwò Musa è fo li o.",
             glosses = "garment Musa PRS wash be_clean FOC",
             free_translation = "Musa is washing the GARMENT clean.")
```


<!-- ## Nyam -->
<!-- ### Details -->
<!-- Glottocode: nyam1285 <br>  -->
<!-- Family: Afro-Asiatic <br> -->
<!-- Macroarea: Africa -->

<!-- ### Coding -->
<!-- IU: 1 <br> Grouping: ? <br> Grammatical: 1 -->

<!-- > Citation [@source2001] -->

<!-- ### Examples -->

<!-- (@language1) @source2001, pg, example 27 -->
<!-- ```{r, echo=FALSE} -->
<!-- gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:", -->
<!--              glosses = "downward NAME-DAT shout say-PST", -->
<!--              free_translation = "I shouted it down to Solo:bu.") -->
<!-- ``` -->

<!-- ## Oko -->
<!-- ### Details -->
<!-- Glottocode: okoe1238 <br>  -->
<!-- Family: Volta-Congo <br> -->
<!-- Macroarea: Africa -->

<!-- ### Coding -->
<!-- IU: 1 <br> Grouping: ? <br> Grammatical: 1 -->

<!-- > Citation [@source2001] -->

<!-- ### Examples -->

<!-- (@language1) @source2001, pg, example 27 -->
<!-- ```{r, echo=FALSE} -->
<!-- gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:", -->
<!--              glosses = "downward NAME-DAT shout say-PST", -->
<!--              free_translation = "I shouted it down to Solo:bu.") -->
<!-- ``` -->


## Olo
### Details
Glottocode: oloo1241 <br> 
Family: Nuclear Torricelli <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

> Olo has a construction that contains more than one verb and for which the verbs are not in a grammatical dependency relationship. The verbs take their full morphological marking irrespective of their serial position. The construction is under a single intonation contour. [@staley2007, 37]

### Examples

(@olo1) @staley2007, 37, example 112
```{r}
gloss_example(transliteration = "Wamnei n-e kali tom n-ila n-au wa-iki",
             glosses = "Wamnei 3FEM-go get.3MASC string_bag 3FEM-carry 3FEM-come give-1SG",
             free_translation = "Wamnei went, got the string bag, and carrying it, she came and gave it to me.")
```

(@olo2) @staley2007, 39, example 115a
```{r}
gloss_example(transliteration = "ne kapi yali n-ila-pe n-e uf",
             glosses = "3FEM get.3PL firewood.PL 3FEM-lift-3PL 3FEM-go village",
             free_translation = "She got the firewood, lifted it and went to the village.")
```
