```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lingglosses)
#library(linguisticsdown)
#linguisticsdown::writeIPA()
```


## Waigeo
### Details
Glottocode: waig1244 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Waima'a!
### Details
Glottocode: waim1251 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: 1 <br> Grammatical: 0

### Examples

> Citation Culhane p.c.

(@waimaa1) Aiton 2016:200, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```



## Wambayan
### Details
Glottocode: wamb1258 <br> 
Family: Mirndi <br>
Macroarea: Australia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 0

> Examples of complex predicates ... are more complicated. In these examples, all of which are in the non-future tense, it is not possible to clearly identify one verb as being the main verb, and the other as a modifier. In fact, it may be that the verbs in these examples are serialised. ...  These examples appear to have the intonational contour of single clauses and thus, are not analysable as coordinate clauses along the lines of those discussed in 8.2.2. However, at this stage of this investigation, such a distinction is based on purely impressionistic grounds...
[@nordlinger1998, 194-195]

### Examples

(@wambaya1) @nordlinger1998, 194, example 7-118
```{r, echo=FALSE}
gloss_example(transliteration = "Gayini g-a yarru ginkanyi nanganangali?",
             glosses = "what.I(NOM) 3SG.SBJ-PST go this_way sneak_away",
             free_translation = "Who was it that snuck off this way?")
```

(@wambaya2) @nordlinger1998, 195, example 7-124
```{r, echo=FALSE}
gloss_example(transliteration = "Yugu-yugu-lumi gin-a irra bard-babu.",
             glosses = "cry-RDP-CAUS 3SG.MASC.A-PST 3PL.ACC run-APPL",
             free_translation = "He took them away (from their mothers) and made them cry.")
```



## Wandala
### Details
Glottocode: wand1278 <br> 
Family: Afro-Asiatic <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## West Coast Bajau
### Details
Glottocode: west2560 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## West Subanon
### Details
Glottocode: west2557 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## West Uvean
### Details
Glottocode: west2516 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Western Bru
### Details
Glottocode: west2397 <br> 
Family: Austroasiatic <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Western Magar
### Details
Glottocode: west218 <br> 
Family: Trans-Himalayan  <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

> Serialized verbs are two juxtaposed verbs which share derivational and inflectional morphology. The first verb in the sequence is a bare stem and the second is inflected, as seen in (@wmagar1). ... Serial verbs, combining two verbs, might be considered parts of separate clauses. Nevertheless, they are treated as single complex verbs, because they are unary in meaning, in other words, they express a single event/ predication. They also share a single argument and a single set of inflectional and derivational morphology, thus are considered to be one syntactic unit. [@grunow-harsta2008, 384]

### Examples

(@wmagar1) @grunow-harsta2008, 384, example 6b
```{r, echo=FALSE}
#gloss_example(transliteration = "jaja-ko ŋɦet-ke ma-cip-raɦ-a",
#             glosses = "child-PL cow-DAT squeeze-come-PST",
#             free_translation = "The children did not come milk the cow.")
```


## Western Tamang
### Details
Glottocode: west2415 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Western Xiangxi Miao
### Details
Glottocode: west2430 <br> 
Family: Hmong-Mien <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

> The fact that [multiverbal constructions] involve multiple verbs without any overt markers of coordination or subordination might be rather incidental, and the author suspects that in at least some cases there might well be more semantic and grammatical similarities between a particular type of multiverbal construction and a particular overtly marked coordinative or subordinative construction than there are between two different types of multiverbal construction. [@sposato2015, 542]

### Examples

(@xong1) @sposato2015, 541, example 12.1
```{r, echo=FALSE}
gloss_example(transliteration = "Aod-ngonl naus beux daos geud giab nonx.",
             glosses = "one-CLF.ANIM bird hit die hold stir_fry eat",
             free_translation = "(I) killed the bird and cooked it (to eat).")
```

(@xong2) @sposato2015, 566, example 12.91
```{r, echo=FALSE}
gloss_example(transliteration = "Yul hnef monx jaont monx ghad dionb nzhaod lol.",
             glosses = "again CLF.day 2SG herd 2SG then lead return come",
             free_translation = "Next time when you go herding, lead the bird back here with you.")
```

## Wooi
### Details
Glottocode: woii1237 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: 0 <br> Grammatical: 0

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Yabêm
### Details
Glottocode: yabe1254 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Yami
### Details
Glottocode: yami1254 <br> 
Family: Austronesian <br>
Macroarea: Papunesia


### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Yimas
### Details
Glottocode: yima1243 <br> 
Family: Lower Sepik <br>
Macroarea: Papunesia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```


## Yongning Na
### Details
Glottocode: yong1270 <br> 
Family: Trans-Himalayan <br>
Macroarea: Eurasia

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

> Most concatenations of verbs in Na fall into one of the following categories: verb phrases which belong to separate clauses but appear adjacently because their nominal arguments are not overt; a complement taking predicate with an adjacent infinitive-like verb; a verb with an aspect marker which has grammaticalized from a verb and thus has an isomorphic verb counterpart; or a resultative/cause-and-effect SVC. ... Symmetrical serial verb constructions tend to match Aktionsart marking on each verb. [@lidz2010, 396]

### Examples

(@yongningna1) @lidz2010, 397, example 566
```{r, echo=FALSE, results='asis'}
gloss_example(transliteration = "tʰi13 lə33- wæ33 zɔ33 tʰɯ33 ʐɯ33mi33 tʰɯ33 lɯ33 lə33- wɔ13 lə33- sɩ31 lə33- yɔ33 hɯ33",
             glosses = "so ACCOMP- call CSM this hearth.room this CLS ACCOMP- return ACCOMP- live ACCOMP- come FUT.REM",
             free_translation = "Having called it, the hearth room will come back to life.")
```





## Yoruba
### Details
Glottocode: yoru1245 <br> 
Family: Volta-Congo <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> ... there are a number of tests that prove the difference between cordination and the morphosyntactically more coherent SVC. One of them is extraction. While it is possible to move *ìwé* 'book' to the clausal-initial position in an SVC (@yoruba2), this is ungrammatical with coordination (@yoruba4).
[@bisang2009, 796]

### Examples

(@yoruba1) @stahlke1970, 61, example 1
```{r, echo=FALSE}
gloss_example(transliteration = "Mo mú ìwé wá ilé.",
             glosses = "1SG take book come house",
             free_translation = "I brought a book home.")
```

(@yoruba2) @bisang2009, 796, example 3'
```{r, echo=FALSE}
gloss_example(transliteration = "Ìwé ni mo mú wá ilé.",
             glosses = "book FOC 1SG take come house",
             free_translation = "It was a book I brought home.")
```

(@yoruba3) @stahlke1970, 78, example 50
```{r, echo=FALSE}
gloss_example(transliteration = "Mo mú ìwé, mo sì wá ilé.",
             glosses = "1SG take book 1SG and come home",
             free_translation = "I picked up a book and came home.")
```

(@yoruba4) @stahlke1970, 79, example 57
```{r, echo=FALSE}
gloss_example(transliteration = "Ìwé ni mo mú, mo sì wá ilé.",
             glosses = "book FOC 1SG take 1SG and come house",
             free_translation = "It was a book I took and I came home.",
             grammaticality = "*")
```

## Yucatec Maya
### Details
Glottocode: yuca1254 <br> 
Family: Mayan <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

> Infinitives do not exist in the Mayan language in the same sense that they do exist in English. ... For the transitive and intransitive conjugations there are ways of putting two or more verbs together, as will be shown in the following sections. ... <br>
Transitive infinitives: the transitive infinitive is formed by putting the acting verb in the intransitive conjugation with whatever tense and person needed and then adding on what would be the infinitive in English. This “infinitive” appears as the fourth form of the transitive conjugation preceded by a Set A pronoun which agrees in person with the person of the acting verb. ... <br>
Intransitive infinitives: the intransitive infinitive is formed by putting the acting verb in the intransitive conjugation with whatever tense and person needed and then adding on what would be the infinitive in English. This “infinitive” appears as the general form of the intransitive conjugation. No personal pronoun accompanies it.[@bollesbolles2014,112-113]

### Examples

Note that the glosses were not originally provided by the authors of the text.

(@ymaya1) @bollesbolles2014, 113, section 108
```{r, echo=FALSE}
gloss_example(transliteration = "Yan in hokol in caxte in uooch.",
             glosses = "have_to 1SG go_out 1SG gather 1SG ",
             free_translation = "I have to go out to gather my food.")
```

(@ymaya2) @bollesbolles2014, 113, section 109
```{r, echo=FALSE}
gloss_example(transliteration = "Yan in hokol okot.",
             glosses = "have_to 1SG go_out dance",
             free_translation = "I have to go out to dance.")
```

## Yuchi
### Details
Glottocode: yuch1247 <br> 
Family: Isolate <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: ?

### Examples

(@yuchi1) @linn2001, 282, example 20b
```{r, echo=FALSE}
#gloss_example(transliteration = "William dzothishi wenõ hõ'nẽ hõ-wethla jẽ",
#             glosses = "William doctor CL/DET(NE) 3SG(EM).ACT-see 3SG.ACT-go PST",
#             free_translation = "William went to see the doctor.")
```


## Zaniza Zapotec
### Details
Glottocode: zani1235 <br> 
Family: Otomanguean <br>
Macroarea: North America

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

## Zayse-Zergulla
### Details
Glottocode: zays1235 <br> 
Family: Ta-Ne-Omotic <br>
Macroarea: Africa

### Coding
IU: 1 <br> Grouping: ? <br> Grammatical: 1

> Citation [@source2001]

### Examples

(@language1) @source2001, pg, example 27
```{r, echo=FALSE}
gloss_example(transliteration = "ko:dulu sulo:bo-mo: wElE sE:-ja:",
             glosses = "downward NAME-DAT shout say-PST",
             free_translation = "I shouted it down to Solo:bu.")
```

